---
title: "RISG_Larv_Analysis"
author: "Amy Zyck"
date: '2024-12-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r include=TRUE}
library(reshape2)
library(ggplot2)
library(tidyr)
library(dplyr)
library(qvalue)
#library(plyr)
#library(stringr)
library(pcadapt)
library(poolSeq)
library(ACER)
library(ggman)
```

```{r}
sessionInfo()
```

## Download Popoolation2 scripts
```{bash, eval = FALSE}
git clone https://github.com/ToBoDev/assessPool.git
```

### Visualize across 10,000 random loci

### Block 2 

```{bash, eval=FALSE}
mawk '!/CHR/' RISG.Block2.cov.sync | cut -f 1-13 | shuf | shuf | head -q -n 11000 > input.b2.rand.sync 

perl ./assessPool/scripts/p2/snp-frequency-diff.pl --input /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/input.b2.rand.sync --output-prefix B2.rand --max-coverage 50000 

mawk -f ./scripts/polarize_freqs B2.rand_rc | grep -Ei '(A{10}|C{10}|G{10}|T{10})'| head -q -n 10000| cut -f10-19 | mawk '!/maa/'   > B2.rand.pool
```

```{r}
pool.data.B2 <- read.table("B2.rand.pool")
df.pool.B2 <- apply(pool.data.B2, c(1, 2), function(x) eval(parse(text = x)))
df_filtered <- df.pool.B2[!apply(df.pool.B2, 1, function(row) all(row == row[1])), ]

pool.data2.B2 <- t(df_filtered)
filename.B2 <- read.pcadapt(pool.data2.B2, type = "pool")

res.B2.rand <- pcadapt(filename.B2, K =5,min.maf = 0.01)

poplist.names <- c(rep("TL0", 4),rep("CB", 3),rep("HB", 3))
p1.B2.rand.df <- data.frame(res.B2.rand$scores)
colnames(p1.B2.rand.df) <- c("PC1","PC2", "PC3", "PC4")
p1.B2.rand.df$POP <- poplist.names
ggplot(p1.B2.rand.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("#56B4E9", "orange","purple"),name="Treatment")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))
```

### Block 3 

```{bash, eval=FALSE}
mawk '!/CHR/' RISG.Block3.cov.sync | cut -f 1-13 | shuf | shuf | head -q -n 11000 > input.b3.rand.sync 

perl ./assessPool/scripts/p2/snp-frequency-diff.pl --input /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/input.b3.rand.sync --output-prefix b3.rand --max-coverage 50000 

mawk -f ./scripts/polarize_freqs b3.rand_rc | grep -Ei '(A{10}|C{10}|G{10}|T{10})'| head -q -n 10000| cut -f10-19 | mawk '!/maa/'   > b3.rand.pool
```

```{r}
pool.data.b3 <- read.table("b3.rand.pool")
df.pool.b3 <- apply(pool.data.b3, c(1, 2), function(x) eval(parse(text = x)))
df_filtered <- df.pool.b3[!apply(df.pool.b3, 1, function(row) all(row == row[1])), ]

pool.data2.b3 <- t(df_filtered)
filename.b3 <- read.pcadapt(pool.data2.b3, type = "pool")

res.b3.rand <- pcadapt(filename.b3, K =5,min.maf = 0.01)

poplist.names <- c(rep("TL0", 4),rep("CB", 3),rep("HB", 3))
p1.b3.rand.df <- data.frame(res.b3.rand$scores)
colnames(p1.b3.rand.df) <- c("PC1","PC2", "PC3", "PC4")
p1.b3.rand.df$POP <- poplist.names
ggplot(p1.b3.rand.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("#56B4E9", "orange","purple"),name="Treatment")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))
```

### Block 4 

```{bash, eval=FALSE}
mawk '!/CHR/' RISG.Block4.cov.sync | cut -f 1-13 | shuf | shuf | head -q -n 11000 > input.b4.rand.sync 

perl ./assessPool/scripts/p2/snp-frequency-diff.pl --input /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/input.b4.rand.sync --output-prefix b4.rand --max-coverage 50000 

mawk -f ./scripts/polarize_freqs b4.rand_rc | grep -Ei '(A{10}|C{10}|G{10}|T{10})'| head -q -n 10000| cut -f10-19 | mawk '!/maa/'   > b4.rand.pool
```

```{r}
pool.data.b4 <- read.table("b4.rand.pool")
df.pool.b4 <- apply(pool.data.b4, c(1, 2), function(x) eval(parse(text = x)))
df_filtered <- df.pool.b4[!apply(df.pool.b4, 1, function(row) all(row == row[1])), ]

pool.data2.b4 <- t(df_filtered)
filename.b4 <- read.pcadapt(pool.data2.b4, type = "pool")

res.b4.rand <- pcadapt(filename.b4, K =5,min.maf = 0.01)

poplist.names <- c(rep("TL0", 4),rep("CB", 3),rep("HB", 3))
p1.b4.rand.df <- data.frame(res.b4.rand$scores)
colnames(p1.b4.rand.df) <- c("PC1","PC2", "PC3", "PC4")
p1.b4.rand.df$POP <- poplist.names
ggplot(p1.b4.rand.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("#56B4E9", "orange","purple"),name="Treatment")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))
```


### All blocks

```{bash, eval=FALSE}
mawk '!/CHR/' RISG.All.Blocks.cov.sync | mawk '$34 > 10' | cut -f 1-33 | shuf -n 50000 | shuf -n 11000  > input.all.rand.sync

perl ./assessPool/scripts/p2/snp-frequency-diff.pl --input input.all.rand.sync --output-prefix all.rand --max-coverage 50000 

mawk -f ./scripts/polarize_freqs all.rand_rc | grep -Ei '(A{20}|C{20}|G{20}|T{20})' -m 10000 | cut -f10-39  > all.rand.pool
```

```{r}
pool.data.all <- read.table("all.rand.pool")
df.pool.all <- apply(pool.data.all, c(1, 2), function(x) eval(parse(text = x)))
df_filtered.all <- df.pool.all[!apply(df.pool.all, 1, function(row) all(row == row[1])), ]

pool.data2.all <- t(df_filtered.all)
filename.all <- read.pcadapt(pool.data2.all, type = "pool")

res.all.rand <- pcadapt(filename.all, K =5,min.maf = 0.0001)

poplist.names <- c(rep("B2", 10),rep("B3", 10),rep("B4", 10))
#poplist.names <- c("B12","B11","B12","B2","B11","B2","B11","B12","B2","B12","B11","B2","B11","B12","B12","B2","B12","B11","B2","B11","B11","B12","B2","B11","B12","B11","B2","B12","B2","B11","B12","B11","B12","B2","B11","B2","B11","B2","B11","B2","B11","B12","B12","B12","B12","B11","B12","B2","B2","B11","B11","B2","B12","B12","B11","B12")
p1.all.rand.df <- data.frame(res.all.rand$scores)
colnames(p1.all.rand.df) <- c("PC1","PC2", "PC3", "PC4")
p1.all.rand.df$POP <- poplist.names
ggplot(p1.all.rand.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("green","yellow","blue"),name="Block")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))
```

### All initial samples

```{bash,eval=FALSE}
mawk -f ./scripts/polarize_freqs all.rand_rc | grep -Ei '(A{20}|C{20}|G{20}|T{20})' -m 10000 | cut -f10-13,20-23,30-33  > all.T0.rand.pool
```

```{r}
pool.data.all.is <- read.table("all.T0.rand.pool")
df.pool.all.is <- apply(pool.data.all.is, c(1, 2), function(x) eval(parse(text = x)))
df_filtered.all.is <- df.pool.all.is[!apply(df.pool.all.is, 1, function(row) all(row == row[1])), ]

pool.data2.all.is <- t(df_filtered.all.is)
filename.all.is <- read.pcadapt(pool.data2.all.is, type = "pool")

res.all.is.rand <- pcadapt(filename.all.is, K =5,min.maf = 0.0001)

poplist.names <- c(rep("B2", 4),rep("B3", 4),rep("B4", 4))
p1.all.rand.df <- data.frame(res.all.is.rand$scores)
colnames(p1.all.rand.df) <- c("PC1","PC2", "PC3", "PC4")
p1.all.rand.df$POP <- poplist.names
ggplot(p1.all.rand.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("green","yellow","blue"),name="Block")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))
```

```{bash, eval=FALSE}
curl -L -O https://raw.githubusercontent.com/jpuritz/Puritz_etal_CASE/refs/heads/main/scripts/add_cov_sync_IS
chmod +x add_cov_sync_IS
```

### Block 2

```{bash, eval=FALSE}
bash ./scripts/sum.sh <(cut -f1,2,3,4-7 RISG.Block2.cov.sync) > RISG.Block2.TL0.sync
paste RISG.Block2.TL0.sync <(cut -f4 RISG.Block2.TL0.sync) <(cut -f4 RISG.Block2.TL0.sync)  > RISG.Block2.TL0sum3.sync

AV_COV=$(mawk -f ./scripts/add_cov_sync <(cut -f4-7 --complement RISG.Block2.cov.sync) | mawk '{sum=sum+$15} END {print sum/NR}')

echo $AV_COV

mawk -f ./scripts/add_cov_sync RISG.Block2.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' | mawk '!/CHR/' | cut -f1-6 > RISG.Block2.TL0sum3nh.sync

mawk -f ./scripts/add_cov_sync_IS RISG.Block2.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' > RISG.Block2.TL0sum3nh.cov.sync

paste RISG.Block2.cov.sync <(mawk -f ./scripts/add_cov_sync RISG.Block2.TL0sum3.sync) | mawk -v x=$AV_COV '$23 >= x' | cut -f1-16 > temp.sync

mv temp.sync RISG.dp20.Block2.COV.sync

#perl ./assessPool/scripts/p2/subsample-synchronized.pl --input RISG.Block2.TL0sum3nh.sync --output RISG.Block2.TL0s.sync --target-coverage 48 --method withoutreplace --max-coverage 10000

mawk -f ./scripts/add_cov_sync <(cut -f4-7,14- --complement RISG.dp20.Block2.COV.sync) > temp.cov

paste temp.cov <(cut -f7 RISG.Block2.TL0sum3nh.cov.sync) | mawk -v OFS='\t' '{if ($12 > $13) {$12=$13; print $0} else {print $0}}' | cut -f1-12> RISG.dp20.Block2.COV
```

```{bash, eval=FALSE}
python ./scripts/sub_sample.py RISG.Block2.TL0sum3nh.cov.sync RISG.Block2.TL0sum3nh.sync RISG.Block2.TL0s.sync

cat <(echo -e "CHROM\tPOS\tREF\tTL0_RS1\tTL0_RS2\tTL0_RS3") RISG.Block2.TL0s.sync > RISG.Block2.TL0sum.sync

paste <(cut -f1,2,3,4 RISG.Block2.TL0sum.sync) <(cut -f8 RISG.dp20.Block2.COV.sync) <(cut -f5 RISG.Block2.TL0sum.sync) <(cut -f9 RISG.dp20.Block2.COV.sync) <(cut -f6 RISG.Block2.TL0sum.sync) <(cut -f10 RISG.dp20.Block2.COV.sync)> Con.RISG.B2.sync

head -1 Con.RISG.B2.sync

mawk '!/CHR/' Con.RISG.B2.sync > Con.RISG.B2.input
tail -n +2 Con.RISG.B2.sync | mawk '!/0:0:0:0:0:0/' > Con2.RISG.B2.input

paste <(cut -f1,2,3,4 RISG.Block2.TL0sum.sync) <(cut -f11 RISG.dp20.Block2.COV.sync) <(cut -f5 RISG.Block2.TL0sum.sync) <(cut -f12 RISG.dp20.Block2.COV.sync) <(cut -f6 RISG.Block2.TL0sum.sync) <(cut -f13 RISG.dp20.Block2.COV.sync)> Hi.RISG.B2.sync

head -1 Hi.RISG.B2.sync
tail -n +2 Hi.RISG.B2.sync > Hi.RISG.B2.input
```


### Block 3

```{bash, eval=FALSE}
bash ./scripts/sum.sh <(cut -f1,2,3,4-7 RISG.Block3.cov.sync) > RISG.Block3.TL0.sync
paste RISG.Block3.TL0.sync <(cut -f4 RISG.Block3.TL0.sync) <(cut -f4 RISG.Block3.TL0.sync)  > RISG.Block3.TL0sum3.sync

AV_COV=$(mawk -f ./scripts/add_cov_sync <(cut -f4-7 --complement RISG.Block3.cov.sync) | mawk '{sum=sum+$15} END {print sum/NR}')

echo $AV_COV

mawk -f ./scripts/add_cov_sync RISG.Block3.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' | mawk '!/CHR/' | cut -f1-6 > RISG.Block3.TL0sum3nh.sync

mawk -f ./scripts/add_cov_sync_IS RISG.Block3.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' > RISG.Block3.TL0sum3nh.cov.sync

paste RISG.Block3.cov.sync <(mawk -f ./scripts/add_cov_sync RISG.Block3.TL0sum3.sync) | mawk -v x=$AV_COV '$23 >= x' | cut -f1-16 > temp.sync

mv temp.sync RISG.dp20.Block3.COV.sync

#perl ./assessPool/scripts/p2/subsample-synchronized.pl --input RISG.Block3.TL0sum3nh.sync --output RISG.Block3.TL0s.sync --target-coverage 48 --method withoutreplace --max-coverage 10000

mawk -f ./scripts/add_cov_sync <(cut -f4-7,14- --complement RISG.dp20.Block3.COV.sync) > temp.cov

paste temp.cov <(cut -f7 RISG.Block3.TL0sum3nh.cov.sync) | mawk -v OFS='\t' '{if ($12 > $13) {$12=$13; print $0} else {print $0}}' | cut -f1-12> RISG.dp20.Block3.COV
```

```{bash, eval=FALSE}
python ./scripts/sub_sample.py RISG.Block3.TL0sum3nh.cov.sync RISG.Block3.TL0sum3nh.sync RISG.Block3.TL0s.sync

cat <(echo -e "CHROM\tPOS\tREF\tTL0_RS1\tTL0_RS2\tTL0_RS3") RISG.Block3.TL0s.sync > RISG.Block3.TL0sum.sync

paste <(cut -f1,2,3,4 RISG.Block3.TL0sum.sync) <(cut -f8 RISG.dp20.Block3.COV.sync) <(cut -f5 RISG.Block3.TL0sum.sync) <(cut -f9 RISG.dp20.Block3.COV.sync) <(cut -f6 RISG.Block3.TL0sum.sync) <(cut -f10 RISG.dp20.Block3.COV.sync)> Con.RISG.B3.sync

head -1 Con.RISG.B3.sync

mawk '!/CHR/' Con.RISG.B3.sync > Con.RISG.B3.input
tail -n +2 Con.RISG.B3.sync | mawk '!/0:0:0:0:0:0/' > Con2.RISG.B3.input

paste <(cut -f1,2,3,4 RISG.Block3.TL0sum.sync) <(cut -f11 RISG.dp20.Block3.COV.sync) <(cut -f5 RISG.Block3.TL0sum.sync) <(cut -f12 RISG.dp20.Block3.COV.sync) <(cut -f6 RISG.Block3.TL0sum.sync) <(cut -f13 RISG.dp20.Block3.COV.sync)> Hi.RISG.B3.sync

head -1 Hi.RISG.B3.sync
tail -n +2 Hi.RISG.B3.sync > Hi.RISG.B3.input
```


### Block 4

```{bash, eval=FALSE}
bash ./scripts/sum.sh <(cut -f1,2,3,4-7 RISG.Block4.cov.sync) > RISG.Block4.TL0.sync
paste RISG.Block4.TL0.sync <(cut -f4 RISG.Block4.TL0.sync) <(cut -f4 RISG.Block4.TL0.sync)  > RISG.Block4.TL0sum3.sync

AV_COV=$(mawk -f ./scripts/add_cov_sync <(cut -f4-7 --complement RISG.Block4.cov.sync) | mawk '{sum=sum+$15} END {print sum/NR}')

echo $AV_COV

mawk -f ./scripts/add_cov_sync RISG.Block4.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' | mawk '!/CHR/' | cut -f1-6 > RISG.Block4.TL0sum3nh.sync

mawk -f ./scripts/add_cov_sync_IS RISG.Block4.TL0sum3.sync | mawk -v x=$AV_COV '$7 >= x' > RISG.Block4.TL0sum3nh.cov.sync

paste RISG.Block4.cov.sync <(mawk -f ./scripts/add_cov_sync RISG.Block4.TL0sum3.sync) | mawk -v x=$AV_COV '$23 >= x' | cut -f1-16 > temp.sync

mv temp.sync RISG.dp20.Block4.COV.sync

#perl ./assessPool/scripts/p2/subsample-synchronized.pl --input RISG.Block4.TL0sum3nh.sync --output RISG.Block4.TL0s.sync --target-coverage 48 --method withoutreplace --max-coverage 10000

mawk -f ./scripts/add_cov_sync <(cut -f4-7,14- --complement RISG.dp20.Block4.COV.sync) > temp.cov

paste temp.cov <(cut -f7 RISG.Block4.TL0sum3nh.cov.sync) | mawk -v OFS='\t' '{if ($12 > $13) {$12=$13; print $0} else {print $0}}' | cut -f1-12> RISG.dp20.Block4.COV
```

```{bash, eval=FALSE}
python ./scripts/sub_sample.py RISG.Block4.TL0sum3nh.cov.sync RISG.Block4.TL0sum3nh.sync RISG.Block4.TL0s.sync

cat <(echo -e "CHROM\tPOS\tREF\tTL0_RS1\tTL0_RS2\tTL0_RS3") RISG.Block4.TL0s.sync > RISG.Block4.TL0sum.sync

paste <(cut -f1,2,3,4 RISG.Block4.TL0sum.sync) <(cut -f8 RISG.dp20.Block4.COV.sync) <(cut -f5 RISG.Block4.TL0sum.sync) <(cut -f9 RISG.dp20.Block4.COV.sync) <(cut -f6 RISG.Block4.TL0sum.sync) <(cut -f10 RISG.dp20.Block4.COV.sync)> Con.RISG.B4.sync

head -1 Con.RISG.B4.sync

mawk '!/CHR/' Con.RISG.B4.sync > Con.RISG.B4.input
tail -n +2 Con.RISG.B4.sync | mawk '!/0:0:0:0:0:0/' > Con2.RISG.B4.input

paste <(cut -f1,2,3,4 RISG.Block4.TL0sum.sync) <(cut -f11 RISG.dp20.Block4.COV.sync) <(cut -f5 RISG.Block4.TL0sum.sync) <(cut -f12 RISG.dp20.Block4.COV.sync) <(cut -f6 RISG.Block4.TL0sum.sync) <(cut -f13 RISG.dp20.Block4.COV.sync)> Hi.RISG.B4.sync

head -1 Hi.RISG.B4.sync
tail -n +2 Hi.RISG.B4.sync > Hi.RISG.B4.input
```


## Outlier identification

### Calculate P-values

```{r}
Qvalue_convert <- function(table) {


table <- spread(table,GROUP,PVAL)  
  
table$CHR <- table$CHROM

table %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035780.1", "1")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035781.1", "2")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035782.1", "3")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035783.1", "4")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035784.1", "5")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035785.1", "6")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035786.1", "7")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035787.1", "8")) %>%
  mutate(CHR = stringr::str_replace(CHR, "NC_035788.1", "9")) %>% 
  mutate(CHR = stringr::str_replace(CHR, "NC_035789.1", "10"))  -> pp1

pp1 <- unite(pp1, "SNP", c("CHROM","BP"),remove = FALSE)

pp1$CHR <- as.numeric(pp1$CHR)  

# Set pi0=1 to equal Benjamini Hochberg
pp1$QCon <- qvalue(pp1$PCon, pi0 = 1)$qvalues
pp1$QHi <- qvalue(pp1$PHi, pi0 = 1)$qvalues

return(pp1)
}

# Function to determine significant loci in various outpus
Significat_subset <-function(pv, alpha, alpha2) {
# Add new logical columns based on conditions
  pv$Sig.Hi <- pv$QHi < alpha
  
  ppsig <- subset(pv, QHi < alpha)
  ppsig <- subset(ppsig, QCon > alpha2)
  
  # Print diagnostics
  print(nrow(pv))
  print(nrow(ppsig))
  print(nrow(ppsig)/nrow(pv))
  
  # Return the modified dataframe
  return(ppsig)
}

group_and_average <- function(df) {
  df %>%
    dplyr::group_by(SNP) %>%
    dplyr::summarize(
      # Use a conditional inside across to handle all-NA groups
      across(
        c(PCon, PHi, CHR, QCon, QHi), 
        ~ if (all(is.na(.))) NA else min(., na.rm = TRUE)),
      # Combine logical columns, retain TRUE if any TRUE exists in the group
      Sig.Hi = any(Sig.Hi, na.rm = TRUE),
      CHROM = dplyr::first(CHROM),
      BP = dplyr::first(BP),
      BLOCK = dplyr::first(BLOCK),
      .groups = "drop"  # Ungroup the result
    )
}
```

Using a modified version of the Cochran Mantel Hanzel tests

### Block 2

```{r}
con.RISG.B2.sync <- read.sync(file="Con2.RISG.B2.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B2.cov <- poolSeq::coverage(con.RISG.B2.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B2.af <- af(con.RISG.B2.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

con.RISG.B2.pvals <- adapted.cmh.test(freq=con.RISG.B2.af, coverage=con.RISG.B2.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(con.RISG.B2.af)/2), MeanStart = TRUE, mincov =20)

pcondf <- as.data.table(cbind(row.names(con.RISG.B2.af),con.RISG.B2.pvals))
colnames(pcondf) <- c("Locus","PVAL")
pcondf$GROUP <- "PCon"

hi.RISG.B2.sync <- read.sync(file="Hi.RISG.B2.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B2.cov <- poolSeq::coverage(hi.RISG.B2.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B2.af <- af(hi.RISG.B2.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

hi.RISG.B2.pvals <- adapted.cmh.test(freq=hi.RISG.B2.af, coverage=hi.RISG.B2.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(hi.RISG.B2.af)/2), MeanStart = TRUE, mincov =20)

phidf <- as.data.table(cbind(row.names(hi.RISG.B2.af),hi.RISG.B2.pvals))
colnames(phidf) <- c("Locus","PVAL")
phidf$GROUP <- "PHi"

RISG.B2.pval.table <- bind_rows(pcondf,phidf)
RISG.B2.pval.table <- RISG.B2.pval.table %>%
  separate(Locus, into = c("CHROM", "BP"), sep = "\\.(?=[^.]+$)")
RISG.B2.pval.table$PVAL <- as.numeric(RISG.B2.pval.table$PVAL)
```

### Block 3

```{r}
con.RISG.B3.sync <- read.sync(file="Con2.RISG.B3.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B3.cov <- poolSeq::coverage(con.RISG.B3.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B3.af <- af(con.RISG.B3.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

con.RISG.B3.pvals <- adapted.cmh.test(freq=con.RISG.B3.af, coverage=con.RISG.B3.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(con.RISG.B3.af)/2), MeanStart = TRUE, mincov =20)

pcondf <- as.data.table(cbind(row.names(con.RISG.B3.af),con.RISG.B3.pvals))
colnames(pcondf) <- c("Locus","PVAL")
pcondf$GROUP <- "PCon"

hi.RISG.B3.sync <- read.sync(file="Hi.RISG.B3.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B3.cov <- poolSeq::coverage(hi.RISG.B3.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B3.af <- af(hi.RISG.B3.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

hi.RISG.B3.pvals <- adapted.cmh.test(freq=hi.RISG.B3.af, coverage=hi.RISG.B3.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(hi.RISG.B3.af)/2), MeanStart = TRUE, mincov =20)

phidf <- as.data.table(cbind(row.names(hi.RISG.B3.af),hi.RISG.B3.pvals))
colnames(phidf) <- c("Locus","PVAL")
phidf$GROUP <- "PHi"

RISG.B3.pval.table <- bind_rows(pcondf,phidf)
RISG.B3.pval.table <- RISG.B3.pval.table %>%
  separate(Locus, into = c("CHROM", "BP"), sep = "\\.(?=[^.]+$)")
RISG.B3.pval.table$PVAL <- as.numeric(RISG.B3.pval.table$PVAL)
```


### Block 4

```{r}
con.RISG.B4.sync <- read.sync(file="Con2.RISG.B4.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B4.cov <- poolSeq::coverage(con.RISG.B4.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
con.RISG.B4.af <- af(con.RISG.B4.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

con.RISG.B4.pvals <- adapted.cmh.test(freq=con.RISG.B4.af, coverage=con.RISG.B4.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(con.RISG.B4.af)/2), MeanStart = TRUE, mincov =20)

pcondf <- as.data.table(cbind(row.names(con.RISG.B4.af),con.RISG.B4.pvals))
colnames(pcondf) <- c("Locus","PVAL")
pcondf$GROUP <- "PCon"

hi.RISG.B4.sync <- read.sync(file="Hi.RISG.B4.input", gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B4.cov <- poolSeq::coverage(hi.RISG.B4.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))
hi.RISG.B4.af <- af(hi.RISG.B4.sync,gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3))

hi.RISG.B4.pvals <- adapted.cmh.test(freq=hi.RISG.B4.af, coverage=hi.RISG.B4.cov, Ne=rep(1000, 3), gen=c(0, 1, 0, 1, 0, 1), repl=c(1, 1, 2, 2, 3,3), poolSize=rep(c(40000,10000), ncol(hi.RISG.B4.af)/2), MeanStart = TRUE, mincov =20)

phidf <- as.data.table(cbind(row.names(hi.RISG.B4.af),hi.RISG.B4.pvals))
colnames(phidf) <- c("Locus","PVAL")
phidf$GROUP <- "PHi"

RISG.B4.pval.table <- bind_rows(pcondf,phidf)
RISG.B4.pval.table <- RISG.B4.pval.table %>%
  separate(Locus, into = c("CHROM", "BP"), sep = "\\.(?=[^.]+$)")
RISG.B4.pval.table$PVAL <- as.numeric(RISG.B4.pval.table$PVAL)
```

```{r}
Qvalue_convert(RISG.B2.pval.table) ->B2.pv
Qvalue_convert(RISG.B3.pval.table) ->B3.pv
Qvalue_convert(RISG.B4.pval.table) ->B4.pv
```

```{r}
pp1 <- rbind(B2.pv,B3.pv,B4.pv)
```

```{r}
alpha = 0.1
alpha2 = 0.1


ppsig.B2.1 <- Significat_subset(B2.pv,alpha,alpha2)

ppsig.B3.1 <- Significat_subset(B3.pv,alpha,alpha2)

ppsig.B4.1 <- Significat_subset(B4.pv,alpha,alpha2)
```

```{r}
ppsig.B2.1$BLOCK <- 2
ppsig.B3.1$BLOCK <- 3
ppsig.B4.1$BLOCK <- 4

ppsig3.FDR10 <- bind_rows(ppsig.B2.1,ppsig.B3.1,ppsig.B4.1)

write.table(ppsig3.FDR10, "Sig.Loci.FDR10", sep="\t", row.names = FALSE, quote = FALSE)
```


```{r}
alpha = 0.05
alpha.B3 = alpha * 2
alpha2 = 0.1


ppsig.B2.05 <- Significat_subset(B2.pv,alpha,alpha2)

ppsig.B3.05 <- Significat_subset(B3.pv,alpha.B3,alpha2)

ppsig.B4.05 <- Significat_subset(B4.pv,alpha,alpha2)
```


```{r}
ppsig.B2.05$BLOCK <- 2
ppsig.B3.05$BLOCK <- 3
ppsig.B4.05$BLOCK <- 4

ppsig3.FDR05 <- bind_rows(ppsig.B2.05,ppsig.B3.05,ppsig.B4.05)

write.table(ppsig3.FDR05, "Sig.Loci.FDR05", sep="\t", row.names = FALSE, quote = FALSE)
```


```{r}
alpha = 0.01
alpha.B3 = alpha * 2
alpha2 = 0.1


ppsig.B2.01 <- Significat_subset(B2.pv,alpha,alpha2)

ppsig.B3.01 <- Significat_subset(B3.pv,alpha.B3,alpha2)

ppsig.B4.01 <- Significat_subset(B4.pv,alpha,alpha2)
```


```{r}
conpp1 <- subset(pp1, QCon < alpha)

ppsig.B2.01$BLOCK <- 2
ppsig.B3.01$BLOCK <- 3
ppsig.B4.01$BLOCK <- 4

ppsig3.FDR01 <- bind_rows(ppsig.B2.01,ppsig.B3.01,ppsig.B4.01)

write.table(ppsig3.FDR01, "Sig.Loci.FDR01", sep="\t", row.names = FALSE, quote = FALSE)
write.table(ppsig.B2.01, "Sig.Loci.FDR01.Block2", sep="\t", row.names = FALSE, quote = FALSE)
write.table(ppsig.B3.01, "Sig.Loci.FDR01.Block3", sep="\t", row.names = FALSE, quote = FALSE)
write.table(ppsig.B4.01, "Sig.Loci.FDR01.Block4", sep="\t", row.names = FALSE, quote = FALSE)
```

## Repeating across all blocks

```{bash, eval=FALSE}
# Extract the TL0 samples and add coverage stats
bash ./scripts/sum.sh <(cut -f1,2,3,4,5,6,7 RISG.All.Blocks.cov.sync) > RISG.B2.AB.TL0sum.sync
bash ./scripts/sum.sh <(cut -f1,2,3,14,15,16,17 RISG.All.Blocks.cov.sync) > RISG.B3.AB.TL0sum.sync
bash ./scripts/sum.sh <(cut -f1,2,3,24,25,26,27 RISG.All.Blocks.cov.sync) > RISG.B4.AB.TL0sum.sync

# Pull out the min.cov column from each Block's initial samples and combine with the full file
# Filter each TL0 min.cov column, retaining min coverage greater than 72 and overall min coverage greater than 1. Save the original sample file columns.  
paste RISG.All.Blocks.cov.sync <(mawk -f ./scripts/add_cov_sync RISG.B2.AB.TL0sum.sync | cut -f5 ) <(mawk -f ./scripts/add_cov_sync RISG.B3.AB.TL0sum.sync | cut -f5) <(mawk -f ./scripts/add_cov_sync RISG.B4.AB.TL0sum.sync | cut -f5) | mawk '$37 >72 && $38 > 72 && $39 > 72 && $34 > 1' | cut -f1-36 > RISG.All.Blocks.cov.fil.sync

# Save the initial samples from this newly filtered file
bash ./scripts/sum.sh <(cut -f1,2,3,4,5,6,7 RISG.All.Blocks.cov.fil.sync) > RISG.B2.AB.TL0sum.sync
bash ./scripts/sum.sh <(cut -f1,2,3,14,15,16,17 RISG.All.Blocks.cov.fil.sync) > RISG.B3.AB.TL0sum.sync
bash ./scripts/sum.sh <(cut -f1,2,3,24,25,26,27 RISG.All.Blocks.cov.fil.sync) > RISG.B4.AB.TL0sum.sync

# Save all final samples from this newly filtered file
mawk -f ./scripts/add_cov_sync <(cut -f1-3,8-13 RISG.All.Blocks.cov.fil.sync ) > B2.AB.cov.sync
mawk -f ./scripts/add_cov_sync <(cut -f1-3,18-23 RISG.All.Blocks.cov.fil.sync) > B3.AB.cov.sync
mawk -f ./scripts/add_cov_sync <(cut -f1-3,28-33 RISG.All.Blocks.cov.fil.sync) > B4.AB.cov.sync

# replicate TL0 file three times, one per treatment replicate
paste RISG.B2.AB.TL0sum.sync <(cut -f4 RISG.B2.AB.TL0sum.sync) <(cut -f4 RISG.B2.AB.TL0sum.sync)  > RISG.B2.AB.TL0sum3.sync

paste RISG.B3.AB.TL0sum.sync <(cut -f4 RISG.B3.AB.TL0sum.sync) <(cut -f4 RISG.B3.AB.TL0sum.sync)  > RISG.B3.AB.TL0sum3.sync

paste RISG.B4.AB.TL0sum.sync <(cut -f4 RISG.B4.AB.TL0sum.sync) <(cut -f4 RISG.B4.AB.TL0sum.sync)  > RISG.B4.AB.TL0sum3.sync

mawk -f ./scripts/add_cov_sync_IS RISG.B2.AB.TL0sum3.sync  > RISG.B2.AB.TL0sum3.cov.sync
mawk -f ./scripts/add_cov_sync_IS RISG.B3.AB.TL0sum3.sync  > RISG.B3.AB.TL0sum3.cov.sync
mawk -f ./scripts/add_cov_sync_IS RISG.B4.AB.TL0sum3.sync  > RISG.B4.AB.TL0sum3.cov.sync

paste B2.AB.cov.sync <(cut -f7 RISG.B2.AB.TL0sum3.cov.sync) | mawk -v OFS='\t' '{if (NR > 1 && $12 > $13) {$12=$13; print $0} else {print $0}}' > temp.cov.sync
cut -f1-12 temp.cov.sync > B2.AB.cov.sync

paste B3.AB.cov.sync <(cut -f7 RISG.B3.AB.TL0sum3.cov.sync) | mawk -v OFS='\t' '{if (NR > 1 && $12 > $13) {$12=$13; print $0} else {print $0}}' > temp.cov.sync
cut -f1-12 temp.cov.sync > B3.AB.cov.sync

paste B4.AB.cov.sync <(cut -f7 RISG.B4.AB.TL0sum3.cov.sync) | mawk -v OFS='\t' '{if (NR > 1 && $12 > $13) {$12=$13; print $0} else {print $0}}' > temp.cov.sync
cut -f1-12 temp.cov.sync > B4.AB.cov.sync
rm temp.cov.sync
```


```{bash, eval=FALSE}
python ./scripts/sub_sample.py RISG.B2.AB.TL0sum3.cov.sync <( mawk '!/CHR/' RISG.B2.AB.TL0sum3.sync) RISG.B2.TL0s.AB.sync
python ./scripts/sub_sample.py RISG.B3.AB.TL0sum3.cov.sync <( mawk '!/CHR/' RISG.B3.AB.TL0sum3.sync) RISG.B3.TL0s.AB.sync
python ./scripts/sub_sample.py RISG.B4.AB.TL0sum3.cov.sync <(mawk '!/CHR/' RISG.B4.AB.TL0sum3.sync) RISG.B4.TL0s.AB.sync

wait 

cat <(echo -e "CHROM\tPOS\tREF\tISB2_RS1\tISB2_RS2\tISB2_RS3") RISG.B2.TL0s.AB.sync > RISG.B2.TL0sum.AB.sync
cat <(echo -e "CHROM\tPOS\tREF\tISB3_RS1\tISB3_RS2\tISB3_RS3") RISG.B3.TL0s.AB.sync > RISG.B3.TL0sum.AB.sync
cat <(echo -e "CHROM\tPOS\tREF\tISB4_RS1\tISB4_RS2\tISB4_RS3") RISG.B4.TL0s.AB.sync > RISG.B4.TL0sum.AB.sync
```

```{bash, eval=FALSE}
paste <(cut -f1-4 RISG.B2.TL0sum.AB.sync) <(cut -f8 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B2.TL0sum.AB.sync) <(cut -f9 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B2.TL0sum.AB.sync) <(cut -f10 RISG.All.Blocks.cov.fil.sync) <(cut -f4 RISG.B3.TL0sum.AB.sync) <(cut -f18 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B3.TL0sum.AB.sync) <(cut -f19 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B3.TL0sum.AB.sync) <(cut -f20 RISG.All.Blocks.cov.fil.sync) <(cut -f4 RISG.B4.TL0sum.AB.sync) <(cut -f28 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B4.TL0sum.AB.sync) <(cut -f29 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B4.TL0sum.AB.sync) <(cut -f30 RISG.All.Blocks.cov.fil.sync) > Con.AB.sync

# Removing rows with NAs - found when calculating allele freq.
tail -n +2 Con.AB.sync | mawk '$2 != 5850793' > Con.AB.input

paste <(cut -f1-4 RISG.B2.TL0sum.AB.sync) <(cut -f11 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B2.TL0sum.AB.sync) <(cut -f12 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B2.TL0sum.AB.sync) <(cut -f13 RISG.All.Blocks.cov.fil.sync) <(cut -f4 RISG.B3.TL0sum.AB.sync) <(cut -f21 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B3.TL0sum.AB.sync) <(cut -f22 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B3.TL0sum.AB.sync) <(cut -f23 RISG.All.Blocks.cov.fil.sync) <(cut -f4 RISG.B4.TL0sum.AB.sync) <(cut -f31 RISG.All.Blocks.cov.fil.sync) <(cut -f5 RISG.B4.TL0sum.AB.sync) <(cut -f32 RISG.All.Blocks.cov.fil.sync) <(cut -f6 RISG.B4.TL0sum.AB.sync) <(cut -f33 RISG.All.Blocks.cov.fil.sync) > Hi.AB.sync

tail -n +2 Hi.AB.sync | mawk '$2 != 7310884 && $2 != 25363715' > Hi.AB.input 
```

```{r}
con.ab.sync <- read.sync(file="Con.AB.input", gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ))
con.ab.cov <- poolSeq::coverage(con.ab.sync,gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ))
con.ab.af <- poolSeq::af(con.ab.sync,gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9))

con.ab.pvals <- adapted.cmh.test(freq=con.ab.af, coverage=con.ab.cov, Ne=c(1000,1000,1000,1000,1000,1000,1000,1000,1000), gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1,1,2,2,3,3,4, 4,5,5,6,6,7,7,8,8,9,9), poolSize=rep(c(40000,10000), ncol(con.ab.af)/2), MeanStart = TRUE, mincov =14)

pcondf <- as.data.table(cbind(row.names(con.ab.af),con.ab.pvals))
colnames(pcondf) <- c("Locus","PVAL")
pcondf$GROUP <- "PCon"

hi.ab.sync <- read.sync(file="Hi.AB.input", gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ))
hi.ab.cov <- poolSeq::coverage(hi.ab.sync,gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ))
hi.ab.af <- af(hi.ab.sync,gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ))

hi.ab.pvals <- adapted.cmh.test(freq=hi.ab.af, coverage=hi.ab.cov, Ne=c(1000,1000,1000,1000,1000,1000,1000,1000,1000), gen=c(0, 1, 0, 1, 0, 1, 0, 1,0,1,0,1,0,1,0,1,0,1), repl=c(1, 1, 2, 2, 3, 3, 4 , 4,5,5,6,6,7,7,8,8,9,9 ), poolSize=rep(c(40000,10000), ncol(hi.ab.af)/2), MeanStart = TRUE, mincov =14)

phidf <- as.data.table(cbind(row.names(hi.ab.af),hi.ab.pvals))
colnames(phidf) <- c("Locus","PVAL")
phidf$GROUP <- "PHi"
```

```{r}
AB.pval.table <- bind_rows(pcondf,phidf)
AB.pval.table <- AB.pval.table %>%
  separate(Locus, into = c("CHROM", "BP"), sep = "\\.(?=[^.]+$)")
AB.pval.table$PVAL <- as.numeric(AB.pval.table$PVAL)
```

```{r}
Qvalue_convert(AB.pval.table) -> AB.pv
```

```{r}
alpha = 0.01
alpha2 =0.1

ppsig.AB <- Significat_subset(AB.pv,alpha,alpha2)
```

```{r}
write.table(ppsig.AB, "Sig.Loci.FDR01.AB", sep="\t", row.names = FALSE, quote = FALSE)

conpp.AB <- subset(AB.pv, QCon < 0.001)
all_con <- bind_rows(conpp1,conpp.AB) %>% group_by(SNP) %>% filter(n()>1)

all_con <- subset(all_con, QCon / QHi < 10)

write.table(all_con, "ALL.CON.SNPS", sep="\t", row.names = FALSE, quote = FALSE)

ppsig.AB$BLOCK <- 10
```

```{r}
total.sig <- anti_join(bind_rows(ppsig.AB,ppsig3.FDR01),all_con, by = "SNP")

write.table(total.sig, "Total.Significant.uf.Loci", sep="\t", row.names = FALSE, quote = FALSE)
```

```{bash, eval=FALSE}
cut -f2,3 Total.Significant.uf.Loci | tail -n +2 | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RISG.Total.Significant.uf.loci.bed

cut -f2,3 ALL.CON.SNPS | tail -n +2 | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > ALL.CON.SNPS.loci.bed

bedtools intersect -wb -a  ALL.CON.SNPS.loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > CON.LOCI

bedtools intersect -wb -a ALL.CON.SNPS.loci.bed -b sorted.ref3.0.gene.bed | cut -f4- | bedtools sort | bedtools merge > con.loci.filter.bed

cat <(echo -e "SNP\tCHROM\tBP") <(bedtools intersect -a RISG.Total.Significant.uf.loci.bed -b con.loci.filter.bed | mawk '{print $1"_"$3"\t"$1"\t"$3}') > ToFilter.Con.Loci.txt
```

```{r}
con_filter <- read.table("ToFilter.Con.Loci.txt",header=TRUE)

total.sig <- unique(anti_join(total.sig,con_filter, by = "SNP"))

write.table(total.sig, "Total.Significant.Loci", sep="\t", row.names = FALSE, quote = FALSE)


grouped_total.sig <- group_and_average(total.sig)

write.table(grouped_total.sig, "Total.Significant.Grouped.Loci", sep="\t", row.names = FALSE, quote = FALSE)
```

### Genes for significant loci

```{bash, eval=FALSE}
cut -f2,3 Total.Significant.Loci | tail -n +2 | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RISG.Total.Significant.loci.bed
bedtools slop -b 100 -i RISG.Total.Significant.loci.bed  -g ../reference.fasta.fai  | bedtools merge -i - >  RISG.Total.Significant.loci.intervals.bed

bedtools intersect -wb -a RISG.Total.Significant.loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.loci.TOTAL.RISG.LOC
```

## PCA of candidate loci 

```{bash, eval=FALSE}
bcftools view --threads 40 B2.RISG.vcf.gz -R <(cut -f2,3 Total.Significant.Loci | tail -n +2 | sort | uniq )| mawk '!/\.:\.:\./' > B2.total.outlier.RISG.dp20.vcf &

bcftools view --threads 40 B3.RISG.vcf.gz -R <(cut -f2,3 Total.Significant.Loci | tail -n +2 | sort | uniq )| mawk '!/\.:\.:\./' > B3.total.outlier.RISG.dp20.vcf &

bcftools view --threads 40 B4.RISG.vcf.gz -R <(cut -f2,3 Total.Significant.Loci | tail -n +2 | sort | uniq )| mawk '!/\.:\.:\./' > B4.total.outlier.RISG.dp20.vcf &

bcftools view --threads 40 RISGTRSdp20g75nDNA.FIL.recode.vcf.gz -R <(cut -f2,3 Total.Significant.Loci | tail -n +2 | sort | uniq )| mawk '!/\.:\.:\./' > total.outlier.RISG.dp20.vcf 

wait

python2 VCFtoPopPool.py B2.total.outlier.RISG.dp20.vcf B2.total.outlier.sync
python2 VCFtoPopPool.py B3.total.outlier.RISG.dp20.vcf B3.total.outlier.sync
python2 VCFtoPopPool.py B4.total.outlier.RISG.dp20.vcf B4.total.outlier.sync 

python2 VCFtoPopPool.py total.outlier.RISG.dp20.vcf AB.total.outlier.sync


mawk '!/CHR/' AB.total.outlier.sync > AB.total.outlier.input
mawk '!/CHR/' B2.total.outlier.sync > B2.total.outlier.input
mawk '!/CHR/' B3.total.outlier.sync > B3.total.outlier.input
mawk '!/CHR/' B4.total.outlier.sync > B4.total.outlier.input
```

```{bash, eval=FALSE}
./assessPool/scripts/p2/snp-frequency-diff.pl --input  B2.total.outlier.input --min-count 1 --min-coverage 1 --output-prefix B2.total.outlier --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  B3.total.outlier.input --min-count 1 --min-coverage 1 --output-prefix B3.total.outlier --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  B4.total.outlier.input --min-count 1 --min-coverage 1 --output-prefix B4.total.outlier --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  AB.total.outlier.input --min-count 1 --min-coverage 1 --output-prefix AB.total.outlier --max-coverage 50000 
```

```{bash, eval=FALSE}
mawk -f ./scripts/polarize_freqs AB.total.outlier_rc | cut -f10-39 | mawk '!/maa/' > AB.total.outlier.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B2.total.outlier_rc | cut -f10-19 | mawk '!/maa/' > B2.total.outlier.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B3.total.outlier_rc | cut -f10-19 | mawk '!/maa/' > B3.total.outlier.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B4.total.outlier_rc | cut -f10-19 | mawk '!/maa/' > B4.total.outlier.RISG.dp20.pool
```

```{r}
pool.data <- read.table("AB.total.outlier.RISG.dp20.pool")
pool.data.b2 <- read.table("B2.total.outlier.RISG.dp20.pool")
pool.data.b3 <- read.table("B3.total.outlier.RISG.dp20.pool")
pool.data.b4 <- read.table("B4.total.outlier.RISG.dp20.pool")

df.pool <- apply(pool.data, c(1, 2), function(x) eval(parse(text = x)))
df.pool.b2 <- apply(pool.data.b2, c(1, 2), function(x) eval(parse(text = x)))
df.pool.b3 <- apply(pool.data.b3, c(1, 2), function(x) eval(parse(text = x)))
df.pool.b4 <- apply(pool.data.b4, c(1, 2), function(x) eval(parse(text = x)))

pool.data2 <- t(df.pool)
pool.data2.b2 <- t(df.pool.b2)
pool.data2.b3 <- t(df.pool.b3)
pool.data2.b4 <- t(df.pool.b4)

filename <- read.pcadapt(pool.data2, type = "pool")
res <- pcadapt(filename, min.maf = 0.1)

filename.b2 <- read.pcadapt(pool.data2.b2, type = "pool")
res.b2 <- pcadapt(filename.b2)

filename.b3 <- read.pcadapt(pool.data2.b3, type = "pool")
res.b3 <- pcadapt(filename.b3)

filename.b4 <- read.pcadapt(pool.data2.b4, type = "pool")
res.b4 <- pcadapt(filename.b4)
```

## All significant loci 

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res,option="screeplot")
```

```{r}
res <- pcadapt(filename, K =4,min.maf = 0.001)
poplist.names <- factor(c("Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress","Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress","Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress"), levels = c("Initial","Control","Stress"))
p1 <- plot(res, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve <- res$singular.values^2 / sum(res$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("All_cand_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

```{r}
poplist.names <- c(rep("Block2", 10),rep("Block3", 10),rep("Block4", 10))
p1 <- plot(res, option = "scores", i = 1, j = 2, pop = poplist.names, col = c("#CC79A7", "#009E73", "#0072B2"))

p1 + xlab(paste0("PC1 (", round(pve[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("All_cand_loci_PCA_byBlock.pdf", width = 7, height = 6, units = "in")
```

## Neutral loci 

```{bash, eval=FALSE}
cut -f2,3 Total.Significant.Loci | tail -n +2 | sort | uniq | awk '{print $1"\t"$2-1"\t"$2}' > exclude_positions.bed
```

```{bash, eval=FALSE}
bcftools view --threads 40 B2.RISG.vcf.gz -T ^exclude_positions.bed | mawk '!/\.:\.:\./' > B2.total.neutral.RISG.dp20.vcf

bcftools view --threads 40 B3.RISG.vcf.gz -T ^exclude_positions.bed | mawk '!/\.:\.:\./' > B3.total.neutral.RISG.dp20.vcf

bcftools view --threads 40 B4.RISG.vcf.gz -T ^exclude_positions.bed | mawk '!/\.:\.:\./' > B4.total.neutral.RISG.dp20.vcf

bcftools view --threads 40 RISGTRSdp20g75nDNA.FIL.recode.vcf.gz -T ^exclude_positions.bed | mawk '!/\.:\.:\./' > total.neutral.RISG.dp20.vcf 

wait

python2 VCFtoPopPool.py B2.total.neutral.RISG.dp20.vcf B2.total.neutral.sync
python2 VCFtoPopPool.py B3.total.neutral.RISG.dp20.vcf B3.total.neutral.sync
python2 VCFtoPopPool.py B4.total.neutral.RISG.dp20.vcf B4.total.neutral.sync 

python2 VCFtoPopPool.py total.neutral.RISG.dp20.vcf AB.total.neutral.sync


mawk '!/CHR/' AB.total.neutral.sync > AB.total.neutral.input
mawk '!/CHR/' B2.total.neutral.sync > B2.total.neutral.input
mawk '!/CHR/' B3.total.neutral.sync > B3.total.neutral.input
mawk '!/CHR/' B4.total.neutral.sync > B4.total.neutral.input
```

```{bash, eval=FALSE}
./assessPool/scripts/p2/snp-frequency-diff.pl --input  B2.total.neutral.input --min-count 1 --min-coverage 1 --output-prefix B2.total.neutral --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  B3.total.neutral.input --min-count 1 --min-coverage 1 --output-prefix B3.total.neutral --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  B4.total.neutral.input --min-count 1 --min-coverage 1 --output-prefix B4.total.neutral --max-coverage 50000 &

./assessPool/scripts/p2/snp-frequency-diff.pl --input  AB.total.neutral.input --min-count 1 --min-coverage 1 --output-prefix AB.total.neutral --max-coverage 50000 
```

```{bash, eval=FALSE}
mawk -f ./scripts/polarize_freqs AB.total.neutral_rc | cut -f10-39 | mawk '!/maa/' > AB.total.neutral.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B2.total.neutral_rc | cut -f10-19 | mawk '!/maa/' > B2.total.neutral.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B3.total.neutral_rc | cut -f10-19 | mawk '!/maa/' > B3.total.neutral.RISG.dp20.pool

mawk -f ./scripts/polarize_freqs B4.total.neutral_rc | cut -f10-19 | mawk '!/maa/' > B4.total.neutral.RISG.dp20.pool
```

```{r}
pool.data.neut <- read.table("AB.total.neutral.RISG.dp20.pool")
pool.data.neut.b2 <- read.table("B2.total.neutral.RISG.dp20.pool")
pool.data.neut.b3 <- read.table("B3.total.neutral.RISG.dp20.pool")
pool.data.neut.b4 <- read.table("B4.total.neutral.RISG.dp20.pool")

df.pool.neut <- apply(pool.data.neut, c(1, 2), function(x) eval(parse(text = x)))
df.pool.neut.b2 <- apply(pool.data.neut.b2, c(1, 2), function(x) eval(parse(text = x)))
df.pool.neut.b3 <- apply(pool.data.neut.b3, c(1, 2), function(x) eval(parse(text = x)))
df.pool.neut.b4 <- apply(pool.data.neut.b4, c(1, 2), function(x) eval(parse(text = x)))

pool.data.neut2 <- t(df.pool.neut)
pool.data.neut2.b2 <- t(df.pool.neut.b2)
pool.data.neut2.b3 <- t(df.pool.neut.b3)
pool.data.neut2.b4 <- t(df.pool.neut.b4)

filename.neut <- read.pcadapt(pool.data.neut2, type = "pool")
res.neut <- pcadapt(filename.neut, min.maf = 0.1)

filename.neut.b2 <- read.pcadapt(pool.data.neut2.b2, type = "pool")
res.neut.b2 <- pcadapt(filename.neut.b2)

filename.neut.b3 <- read.pcadapt(pool.data.neut2.b3, type = "pool")
res.neut.b3 <- pcadapt(filename.neut.b3)

filename.neut.b4 <- read.pcadapt(pool.data.neut2.b4, type = "pool")
res.neut.b4 <- pcadapt(filename.neut.b4)
```

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.neut$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.neut,option="screeplot")
```

```{r}
res.neut <- pcadapt(filename.neut, K =4,min.maf = 0.001)
poplist.names <- factor(c("Initial","Initial","Initial","Initial","Control","Control","Control","Stres.neuts","Stres.neuts","Stres.neuts","Initial","Initial","Initial","Initial","Control","Control","Control","Stres.neuts","Stres.neuts","Stres.neuts","Initial","Initial","Initial","Initial","Control","Control","Control","Stres.neuts","Stres.neuts","Stres.neuts"), levels = c("Initial","Control","Stres.neuts"))
p1 <- plot(res.neut, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve <- res.neut$singular.values^2 / sum(res.neut$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("All_neut_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

```{r}
poplist.names <- c(rep("Block2", 10),rep("Block3", 10),rep("Block4", 10))
p1 <- plot(res.neut, option = "scores", i = 1, j = 2, pop = poplist.names, col = c("#CC79A7", "#009E73", "#0072B2"))

p1 + xlab(paste0("PC1 (", round(pve[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("All_neut_loci_PCA_byBlock.pdf", width = 7, height = 6, units = "in")
```


## Block 2 only

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.b2$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.b2,option="screeplot")
```

```{r}
res.b2 <- pcadapt(filename.b2, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.b2, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b2 <- res.b2$singular.values^2 / sum(res.b2$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b2[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b2[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B2_cand_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

```{r}
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
B2.out.df <- data.frame(res.b2$scores)
colnames(B2.out.df) <- c("PC1","PC2")
B2.out.df$POP <- poplist.names
p2 <- ggplot(B2.out.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("purple", "#56B4E9", "orange"), name = "Pooled larval samples")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))

pve.b2 <- res.b2$singular.values^2 / sum(res.b2$singular.values^2)

p2 + xlab(paste0("PC1 (", round(pve.b2[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b2[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B2_cand_loci_PCA_large.pdf", width = 7, height = 7, units = "in")
```

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.neut.b2$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.neut.b2,option="screeplot")
```

```{r}
res.neut.b2 <- pcadapt(filename.neut.b2, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.neut.b2, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b2 <- res.neut.b2$singular.values^2 / sum(res.neut.b2$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b2[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b2[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B2_neut_loci_PCA.pdf", width = 7, height = 6, units = "in")
```


## Block 3 only

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.b3$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.b3,option="screeplot")
```

```{r}
res.b3 <- pcadapt(filename.b3, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.b3, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b3 <- res.b3$singular.values^2 / sum(res.b3$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b3[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b3[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B3_cand_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

```{r}
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
B3.out.df <- data.frame(res.b3$scores)
colnames(B3.out.df) <- c("PC1","PC2")
B3.out.df$POP <- poplist.names
p3 <- ggplot(B3.out.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("purple", "#56B4E9", "orange"), name = "Pooled larval samples")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))

pve.b3 <- res.b3$singular.values^2 / sum(res.b3$singular.values^2)

p3 + xlab(paste0("PC1 (", round(pve.b3[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b3[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B3_cand_loci_PCA_large.pdf", width = 7, height = 7, units = "in")
```

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.neut.b3$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.neut.b3,option="screeplot")
```

```{r}
res.neut.b3 <- pcadapt(filename.neut.b3, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.neut.b3, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b3 <- res.neut.b3$singular.values^2 / sum(res.neut.b3$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b3[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b3[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B3_neut_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

## Block 4 only

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.b4$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.b4,option="screeplot")
```

```{r}
res.b4 <- pcadapt(filename.b4, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.b4, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b4 <- res.b4$singular.values^2 / sum(res.b4$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b4[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b4[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B4_cand_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

```{r}
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
B4.out.df <- data.frame(res.b4$scores)
colnames(B4.out.df) <- c("PC1","PC2")
B4.out.df$POP <- poplist.names
p4 <- ggplot(B4.out.df, aes(x=PC1, y= PC2, fill= POP)) + geom_point(aes(alpha=0.2), size =10,shape=21, col="black")+  scale_fill_manual(values=c("purple", "#56B4E9", "orange"), name = "Pooled larval samples")+ 
guides(alpha=FALSE) +theme(axis.title.x = element_text(size = 24),axis.title.y = element_text(size = 24))

pve.b4 <- res.b4$singular.values^2 / sum(res.b4$singular.values^2)

p4 + xlab(paste0("PC1 (", round(pve.b4[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b4[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B4_cand_loci_PCA_large.pdf", width = 7, height = 7, units = "in")
```

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.neut.b4$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.neut.b4,option="screeplot")
```

```{r}
res.neut.b4 <- pcadapt(filename.neut.b4, K =2,min.maf = 0.001)
poplist.names <- factor(c(rep("Initial", 4),rep("Control", 3),rep("Stress", 3)), levels = c("Initial","Control","Stress"))
p1 <- plot(res.neut.b4, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("purple", "#56B4E9", "orange"))

pve.b4 <- res.neut.b4$singular.values^2 / sum(res.neut.b4$singular.values^2)

p1 + xlab(paste0("PC1 (", round(pve.b4[1]*100, 2), "%)")) +
     ylab(paste0("PC2 (", round(pve.b4[2]*100, 2), "%)")) +
     ggtitle(NULL) +
       theme_bw() +
  theme(panel.grid.minor = element_blank())
ggsave("B4_neut_loci_PCA.pdf", width = 7, height = 6, units = "in")
```

## Pairwise Fst

```{r}
library(poolfstat)
```

### All Blocks

```{r}
##Using sync file with outlier larval full population samples samples

out_AB_samples <- c("B2_TL0_1", "B2_TL0_2", "B2_TL0_3", "B2_TL0_4", "B2_TL1_CB1", "B2_TL1_CB2", "B2_TL1_CB3", "B2_TL1_HB1", "B2_TL1_HB2", "B2_TL1_HB3", "B3_TL0_1", "B3_TL0_2", "B3_TL0_3", "B3_TL0_4", "B3_TL1_CB1", "B3_TL1_CB2", "B3_TL1_CB3", "B3_TL1_HB1", "B3_TL1_HB2", "B3_TL1_HB3", "B4_TL0_1", "B4_TL0_2", "B4_TL0_3", "B4_TL0_4", "B4_TL1_CB1", "B4_TL1_CB2", "B4_TL1_CB3", "B4_TL1_HB1", "B4_TL1_HB2", "B4_TL1_HB3")

out_ABdstbn <- popsync2pooldata("AB.total.outlier.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000, 40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000, 40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = out_AB_samples, min.rc = 1, min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, noindel = TRUE, nlines.per.readblock = 1000)

#Compute Fst for every SNP
out_ABdstbn_SNPFst <- computeFST(out_ABdstbn, method = "Anova")

#Compute pairwise Fst matrix
out_ABdstbn_Fst <- compute.pairwiseFST(out_ABdstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.001, output.snp.values = FALSE)

out_ABdstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(out_ABdstbn_Fst@PairwiseFSTmatrix))

RISG_AB_Fst <- out_ABdstbn_PairwiseFSTMatrix[c("B2_TL0_1", "B2_TL0_2", "B2_TL0_3", "B2_TL0_4", "B2_TL1_CB1", "B2_TL1_CB2", "B2_TL1_CB3", "B2_TL1_HB1", "B2_TL1_HB2", "B2_TL1_HB3", "B3_TL0_1", "B3_TL0_2", "B3_TL0_3", "B3_TL0_4", "B3_TL1_CB1", "B3_TL1_CB2", "B3_TL1_CB3", "B3_TL1_HB1", "B3_TL1_HB2", "B3_TL1_HB3", "B4_TL0_1", "B4_TL0_2", "B4_TL0_3", "B4_TL0_4", "B4_TL1_CB1", "B4_TL1_CB2", "B4_TL1_CB3", "B4_TL1_HB1", "B4_TL1_HB2", "B4_TL1_HB3"), c("B2_TL0_1", "B2_TL0_2", "B2_TL0_3", "B2_TL0_4", "B2_TL1_CB1", "B2_TL1_CB2", "B2_TL1_CB3", "B2_TL1_HB1", "B2_TL1_HB2", "B2_TL1_HB3", "B3_TL0_1", "B3_TL0_2", "B3_TL0_3", "B3_TL0_4", "B3_TL1_CB1", "B3_TL1_CB2", "B3_TL1_CB3", "B3_TL1_HB1", "B3_TL1_HB2", "B3_TL1_HB3", "B4_TL0_1", "B4_TL0_2", "B4_TL0_3", "B4_TL0_4", "B4_TL1_CB1", "B4_TL1_CB2", "B4_TL1_CB3", "B4_TL1_HB1", "B4_TL1_HB2", "B4_TL1_HB3")]
write.csv(RISG_AB_Fst, "RISG_AB_Fst_Outliers.csv")
```


### Block 2

```{r}
##Using sync file with outlier larval full population samples samples

out_B2_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

out_B2dstbn <- popsync2pooldata("B2.total.outlier.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = out_B2_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 100)

#Compute Fst for every SNP
out_B2dstbn_SNPFst <- computeFST(out_B2dstbn, method = "Anova")

#Compute pairwise Fst matrix
out_B2dstbn_Fst <- compute.pairwiseFST(out_B2dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

out_B2dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(out_B2dstbn_Fst@PairwiseFSTmatrix))

RISG_B2_Fst <- out_B2dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B2_Fst, "RISG_B2_Fst_Outliers.csv")
```

```{r}
# Plotting
B2_fst <- read.csv("./RISG_B2_Fst_Outliers.csv")

##Find mean Fst for each treatment replicate (since compared to four initial samples)
# Control Rep1
CB1 = (0.011807966 + 0.015026246 + 0.012726436 + 0.013999954)/4
# Control Rep2
CB2 = (0.01660931 + 0.017818079 + 0.017607014 + 0.01751911)/4
# Control Rep3
CB3 = (0.009793842 + 0.010737179 + 0.01049953 + 0.011150936)/4

# Stress Rep1
HB1 = (0.057619337 + 0.060887444 + 0.057891568 + 0.059826896)/4
# Stress Rep2
HB2 = (0.06186405 + 0.066412052 + 0.062987343 + 0.066231627)/4
# Stress Rep3
HB3 = (0.065465738 + 0.070315896 + 0.065943275 + 0.067477377)/4

# Create the dataframe
df.b2 <- data.frame(
  Sample = c("TL0", "CB1", "CB2", "CB3", "HB1", "HB2", "HB3"),
  Treatment = c("Initial", "Control", "Control", "Control", "Stress", "Stress", "Stress"),
  Fst = c(0, CB1, CB2, CB3, HB1, HB2, HB3)
)

# Print the dataframe
print(df.b2)
```

```{r}
df.b2$Treatment <- factor(df.b2$Treatment, levels = c("Initial", "Control", "Stress"))

p.b2 <- ggplot(df.b2, aes(x = Treatment, y = Fst, colour = Treatment)) +
    geom_point(size = 2.5) +
    ylim(0, 0.07) +
    scale_colour_manual("Treatment", values = c("Initial" = "purple", "Control" = "#56B4E9", "Stress" = "orange")) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "none") 
p.b2

p.b2 <- ggsave("B2_out_fst.pdf", width = 6, height = 7, units = "in")
```

```{r}
##Using sync file with neutral larval B2 samples

neut_B2_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

neut_B2dstbn <- popsync2pooldata("B2.total.neutral.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = neut_B2_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 1000)

#Compute Fst for every SNP
neut_B2dstbn_SNPFst <- computeFST(neut_B2dstbn, method = "Anova")

#Compute pairwise Fst matrix
neut_B2dstbn_Fst <- compute.pairwiseFST(neut_B2dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

neut_B2dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(neut_B2dstbn_Fst@PairwiseFSTmatrix))

RISG_B2_Fst_Neut <- neut_B2dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B2_Fst_Neut, "RISG_B2_Fst_Neutral.csv")
```

### Block 3

```{r}
##Using sync file with outlier larval full population samples samples

out_B3_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

out_B3dstbn <- popsync2pooldata("B3.total.outlier.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = out_B3_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 100)

#Compute Fst for every SNP
out_B3dstbn_SNPFst <- computeFST(out_B3dstbn, method = "Anova")

#Compute pairwise Fst matrix
out_B3dstbn_Fst <- compute.pairwiseFST(out_B3dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

out_B3dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(out_B3dstbn_Fst@PairwiseFSTmatrix))

RISG_B3_Fst <- out_B3dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B3_Fst, "RISG_B3_Fst_Outliers.csv")
```

```{r}
# Plotting
B3_fst <- read.csv("./RISG_B3_Fst_Outliers.csv")

##Find mean Fst for each treatment replicate (since compared to four initial samples)
# Control Rep1
CB1 = (0.017708856 + 0.014380814 + 0.015249293 + 0.013490189)/4
# Control Rep2
CB2 = (0.012909324 + 0.009340816 + 0.012177128 + 0.011700868)/4
# Control Rep3
CB3 = (0.01110532 + 0.009031545 + 0.008387985 + 0.007971423)/4

# Stress Rep1
HB1 = (0.037290022 + 0.035108102 + 0.036709939 + 0.035324972)/4
# Stress Rep2
HB2 = (0.02919536 + 0.028835029 + 0.029710279 + 0.026063932)/4
# Stress Rep3
HB3 = (0.022885231 + 0.021126912 + 0.021246467 + 0.021437351)/4

# Create the dataframe
df.b3 <- data.frame(
  Sample = c("TL0", "CB1", "CB2", "CB3", "HB1", "HB2", "HB3"),
  Treatment = c("Initial", "Control", "Control", "Control", "Stress", "Stress", "Stress"),
  Fst = c(0, CB1, CB2, CB3, HB1, HB2, HB3)
)

# Print the dataframe
print(df.b3)
```

```{r}
df.b3$Treatment <- factor(df.b3$Treatment, levels = c("Initial", "Control", "Stress"))

p.b3 <- ggplot(df.b3, aes(x = Treatment, y = Fst, colour = Treatment)) +
    geom_point(size = 2.5) +
    ylim(0, 0.07) +
    scale_colour_manual("Treatment", values = c("Initial" = "purple", "Control" = "#56B4E9", "Stress" = "orange")) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "none") 
p.b3

p.b3 <- ggsave("B3_out_fst.pdf", width = 6, height = 7, units = "in")
```

```{r}
##Using sync file with neutral larval B3 samples

neut_B3_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

neut_B3dstbn <- popsync2pooldata("B3.total.neutral.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = neut_B3_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 1000)

#Compute Fst for every SNP
neut_B3dstbn_SNPFst <- computeFST(neut_B3dstbn, method = "Anova")

#Compute pairwise Fst matrix
neut_B3dstbn_Fst <- compute.pairwiseFST(neut_B3dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

neut_B3dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(neut_B3dstbn_Fst@PairwiseFSTmatrix))

RISG_B3_Fst_Neut <- neut_B3dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B3_Fst_Neut, "RISG_B3_Fst_Neutral.csv")
```


### Block 4

```{r}
##Using sync file with outlier larval full population samples samples

out_B4_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

out_B4dstbn <- popsync2pooldata("B4.total.outlier.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = out_B4_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 100)

#Compute Fst for every SNP
out_B4dstbn_SNPFst <- computeFST(out_B4dstbn, method = "Anova")

#Compute pairwise Fst matrix
out_B4dstbn_Fst <- compute.pairwiseFST(out_B4dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

out_B4dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(out_B4dstbn_Fst@PairwiseFSTmatrix))

RISG_B4_Fst <- out_B4dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B4_Fst, "RISG_B4_Fst_Outliers.csv")
```

```{r}
# Plotting
B4_fst <- read.csv("./RISG_B4_Fst_Outliers.csv")

##Find mean Fst for each treatment replicate (since compared to four initial samples)
# Control Rep1
CB1 = (0.009442129 + 0.012191104 + 0.012538055 + 0.010455243)/4
# Control Rep2
CB2 = (0.006959645 + 0.009939153 + 0.010391017 + 0.006471974)/4
# Control Rep3
CB3 = (0.007595704 + 0.01005335 + 0.01058412 + 0.011185867)/4

# Stress Rep1
HB1 = (0.032993081 + 0.033843252 + 0.033352225 + 0.035699796)/4
# Stress Rep2
HB2 = (0.034327993 + 0.03943782 + 0.0356146 + 0.036040803)/4
# Stress Rep3
HB3 = (0.038761757 + 0.041929297 + 0.041715464 + 0.041830461)/4

# Create the dataframe
df.b4 <- data.frame(
  Sample = c("TL0", "CB1", "CB2", "CB3", "HB1", "HB2", "HB3"),
  Treatment = c("Initial", "Control", "Control", "Control", "Stress", "Stress", "Stress"),
  Fst = c(0, CB1, CB2, CB3, HB1, HB2, HB3)
)

# Print the dataframe
print(df.b4)
```

```{r}
df.b4$Treatment <- factor(df.b4$Treatment, levels = c("Initial", "Control", "Stress"))

p.b4 <- ggplot(df.b4, aes(x = Treatment, y = Fst, colour = Treatment)) +
    geom_point(size = 2.5) +
    ylim(0, 0.07) +
    scale_colour_manual("Treatment", values = c("Initial" = "purple", "Control" = "#56B4E9", "Stress" = "orange")) +
    theme_bw(base_size = 12) +
    theme(panel.grid.minor = element_blank(), legend.position = "none") 
p.b4

p.b4 <- ggsave("B4_out_fst.pdf", width = 6, height = 7, units = "in")
```

```{r}
##Using sync file with neutral larval B4 samples

neut_B4_samples <- c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")

neut_B4dstbn <- popsync2pooldata("B4.total.neutral.input", poolsizes = c(40000, 40000, 40000, 40000, 10000, 10000, 10000, 10000, 10000, 10000), poolnames = neut_B4_samples, min.rc = 0, min.cov.per.pool = 0, max.cov.per.pool = 50000, min.maf = 0, noindel = TRUE, nlines.per.readblock = 1000)

#Compute Fst for every SNP
neut_B4dstbn_SNPFst <- computeFST(neut_B4dstbn, method = "Anova")

#Compute pairwise Fst matrix
neut_B4dstbn_Fst <- compute.pairwiseFST(neut_B4dstbn, method = "Anova", min.cov.per.pool = 1, max.cov.per.pool = 50000, min.maf = 0.1, output.snp.values = FALSE)

neut_B4dstbn_PairwiseFSTMatrix <- as.matrix(as.data.frame(neut_B4dstbn_Fst@PairwiseFSTmatrix))

RISG_B4_Fst_Neut <- neut_B4dstbn_PairwiseFSTMatrix[c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3"), c("TL0_1", "TL0_2", "TL0_3", "TL0_4", "TL1_CB1", "TL1_CB2", "TL1_CB3", "TL1_HB1", "TL1_HB2", "TL1_HB3")]
write.csv(RISG_B4_Fst_Neut, "RISG_B4_Fst_Neutral.csv")
```


## Global Fst Plots

```{r}
# Extracting Fst values and chromosome and position info for B2 cand loci
out_B2_locus_fst <- as.data.frame(out_B2dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

out_B2_locus_chrom <- as.data.frame(out_B2dstbn@snp.info[["Chromosome"]])

out_B2_locus_pos <- as.data.frame(out_B2dstbn@snp.info[["Position"]])

out_B2_locus_fst_bypos <- bind_cols(out_B2_locus_fst,out_B2_locus_chrom,out_B2_locus_pos)

colnames(out_B2_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers 
out_B2_locus_fst_bypos$B2_sel_outlier <- TRUE
```

```{r}
# Extracting Fst values and chromosome and position info for B2 neutral loci
neut_B2_locus_fst <- as.data.frame(neut_B2dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

neut_B2_locus_chrom <- as.data.frame(neut_B2dstbn@snp.info[["Chromosome"]])

neut_B2_locus_pos <- as.data.frame(neut_B2dstbn@snp.info[["Position"]])

neut_B2_locus_fst_bypos <- bind_cols(neut_B2_locus_fst,neut_B2_locus_chrom,neut_B2_locus_pos)

colnames(neut_B2_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers 
neut_B2_locus_fst_bypos$B2_sel_outlier <- FALSE
```

```{r}
combined_B2_global_Fst <- bind_rows(out_B2_locus_fst_bypos, neut_B2_locus_fst_bypos)
```

```{r}
# Change naming of chromosomes
chr_labs <- c("Chr 1", "Chr 2", "Chr 3", "Chr 4", "Chr 5", "Chr 6", "Chr 7", "Chr 8", "Chr 9","Chr 10")
names(chr_labs) <- c("NC_035780.1", "NC_035781.1", "NC_035782.1", "NC_035783.1", "NC_035784.1", "NC_035785.1", "NC_035786.1", "NC_035787.1", "NC_035788.1", "NC_035789.1")

# Create the plot
p <- combined_B2_global_Fst %>%
  mutate(pos_mb = Position / 1000000) %>%
  ggplot(aes(x = pos_mb, y = Fst)) +
  # Add grey points for non-outliers
  geom_point(color = "grey", size = 1) +
  # Add orange points for outliers
  geom_point(data = . %>% filter(B2_sel_outlier == TRUE), 
             aes(color = "Block 2 candidate loci"), size = 2) +
  facet_wrap(~Chromosome, scales = "free_x", nrow = 1, labeller = labeller(Chromosome = chr_labs)) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) + 
  theme_minimal() +
  scale_color_manual(values = c("Block 2 candidate loci" = "#CC79A7"),
                     name = NULL) +
  labs(x = "Position (Mb)", y = expression(F[ST])) +
  ylim(-0.036, 0.3) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10, face = "bold"),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "right")
ggsave("B2_global_fst.pdf", width = 12, height = 5, units = "in")
# Display the plot
print(p)
```

```{r}
# Extracting Fst values and chromosome and position info for B3 cand loci
out_B3_locus_fst <- as.data.frame(out_B3dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

out_B3_locus_chrom <- as.data.frame(out_B3dstbn@snp.info[["Chromosome"]])

out_B3_locus_pos <- as.data.frame(out_B3dstbn@snp.info[["Position"]])

out_B3_locus_fst_bypos <- bind_cols(out_B3_locus_fst,out_B3_locus_chrom,out_B3_locus_pos)

colnames(out_B3_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers
out_B3_locus_fst_bypos$B3_sel_outlier <- TRUE
```

```{r}
# Extracting Fst values and chromosome and position info for B3 neutral loci
neut_B3_locus_fst <- as.data.frame(neut_B3dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

neut_B3_locus_chrom <- as.data.frame(neut_B3dstbn@snp.info[["Chromosome"]])

neut_B3_locus_pos <- as.data.frame(neut_B3dstbn@snp.info[["Position"]])

neut_B3_locus_fst_bypos <- bind_cols(neut_B3_locus_fst,neut_B3_locus_chrom,neut_B3_locus_pos)

colnames(neut_B3_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers 
neut_B3_locus_fst_bypos$B3_sel_outlier <- FALSE
```

```{r}
combined_B3_global_Fst <- bind_rows(out_B3_locus_fst_bypos, neut_B3_locus_fst_bypos)
```

```{r}
# Change naming of chromosomes
chr_labs <- c("Chr 1", "Chr 2", "Chr 3", "Chr 4", "Chr 5", "Chr 6", "Chr 7", "Chr 8", "Chr 9","Chr 10")
names(chr_labs) <- c("NC_035780.1", "NC_035781.1", "NC_035782.1", "NC_035783.1", "NC_035784.1", "NC_035785.1", "NC_035786.1", "NC_035787.1", "NC_035788.1", "NC_035789.1")

# Create the plot
p <- combined_B3_global_Fst %>%
  mutate(pos_mb = Position / 1000000) %>%
  ggplot(aes(x = pos_mb, y = Fst)) +
  # Add grey points for non-outliers
  geom_point(color = "grey", size = 1) +
  # Add orange points for outliers
  geom_point(data = . %>% filter(B3_sel_outlier == TRUE), 
             aes(color = "Block 3 candidate loci"), size = 2) +
  facet_wrap(~Chromosome, scales = "free_x", nrow = 1, labeller = labeller(Chromosome = chr_labs)) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) + 
  theme_minimal() +
  scale_color_manual(values = c("Block 3 candidate loci" = "#009E73"),
                     name = NULL) +
  labs(x = "Position (Mb)", y = expression(F[ST])) +
  ylim(-0.036, 0.4) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10, face = "bold"),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "right")
ggsave("B3_global_fst.pdf", width = 12, height = 5, units = "in")
# Display the plot
print(p)
```

```{r}
# Extracting Fst values and chromosome and position info for neutral loci
out_B4_locus_fst <- as.data.frame(out_B4dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

out_B4_locus_chrom <- as.data.frame(out_B4dstbn@snp.info[["Chromosome"]])

out_B4_locus_pos <- as.data.frame(out_B4dstbn@snp.info[["Position"]])

out_B4_locus_fst_bypos <- bind_cols(out_B4_locus_fst,out_B4_locus_chrom,out_B4_locus_pos)

colnames(out_B4_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers
out_B4_locus_fst_bypos$B4_sel_outlier <- TRUE
```

```{r}
# Extracting Fst values and chromosome and position info for B4 neutral loci
neut_B4_locus_fst <- as.data.frame(neut_B4dstbn_SNPFst[["snp.Fstats"]][["Fst"]])

neut_B4_locus_chrom <- as.data.frame(neut_B4dstbn@snp.info[["Chromosome"]])

neut_B4_locus_pos <- as.data.frame(neut_B4dstbn@snp.info[["Position"]])

neut_B4_locus_fst_bypos <- bind_cols(neut_B4_locus_fst,neut_B4_locus_chrom,neut_B4_locus_pos)

colnames(neut_B4_locus_fst_bypos) <- c("Fst", "Chromosome", "Position")

# Adding columns to specify loci as outliers 
neut_B4_locus_fst_bypos$B4_sel_outlier <- FALSE
```

```{r}
combined_B4_global_Fst <- bind_rows(out_B4_locus_fst_bypos, neut_B4_locus_fst_bypos)
```

```{r}
# Change naming of chromosomes
chr_labs <- c("Chr 1", "Chr 2", "Chr 3", "Chr 4", "Chr 5", "Chr 6", "Chr 7", "Chr 8", "Chr 9","Chr 10")
names(chr_labs) <- c("NC_035780.1", "NC_035781.1", "NC_035782.1", "NC_035783.1", "NC_035784.1", "NC_035785.1", "NC_035786.1", "NC_035787.1", "NC_035788.1", "NC_035789.1")

# Create the plot
p <- combined_B4_global_Fst %>%
  mutate(pos_mb = Position / 1000000) %>%
  ggplot(aes(x = pos_mb, y = Fst)) +
  # Add grey points for non-outliers
  geom_point(color = "grey", size = 1) +
  # Add orange points for outliers
  geom_point(data = . %>% filter(B4_sel_outlier == TRUE), 
             aes(color = "Block 4 candidate loci"), size = 2) +
  facet_wrap(~Chromosome, scales = "free_x", nrow = 1, labeller = labeller(Chromosome = chr_labs)) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) + 
  theme_minimal() +
  scale_color_manual(values = c("Block 4 candidate loci" = "#0072B2"),
                     name = NULL) +
  labs(x = "Position (Mb)", y = expression(F[ST])) +
  ylim(-0.036, 0.3) +  
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        strip.background = element_blank(),
        strip.text.x = element_text(size = 10, face = "bold"),
        panel.spacing = unit(0.1, "lines"),
        legend.position = "right")
ggsave("B4_global_fst.pdf", width = 12, height = 5, units = "in")
# Display the plot
print(p)
```

## Manhattan Plots

```{r}
library(patchwork)
```

```{r}
b2.h <- subset(total.sig,  BLOCK == "2" & Sig.Hi == TRUE)
b2.c <- subset(B2.pv, QCon < 0.1)


B2.pp <- B2.pv %>%
  mutate(
    Group = case_when(
      SNP %in% b2.c$SNP ~ "Control Treatment Filtered",
      SNP %in% all_con$SNP ~ "Control Treatment Filtered",
      SNP %in% con_filter$SNP ~ "Control Treatment Filtered",
      TRUE ~ "Neutral"
    )
  )

man <-ggman(B2.pp, pvalue="QHi", chr="CHR", snp="SNP",bp="BP", relative.positions = TRUE ,sigLine = NA, pointSize=1.5,title = "", alpha = 0.5)

dfm <- man[[1]]

dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmo <- semi_join(dfm, b2.h, by = "SNP")
dfmo$Group <- "Block 2 Outlier"

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})
b2.Hi <-ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(aes(alpha=0.5, shape=Group, size = marker))+  
  geom_point(data=dfmo, fill="#CC79A7", aes(x=index, y=marker, shape=Group, size=marker),  color = "#CC79A7")+
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = FALSE,alpha=FALSE, fill = FALSE, size = FALSE) +
  scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm$marker+1,na.rm=TRUE)))+
  scale_size_continuous(range = c(0.25,3)) +
  ggtitle("Block A") +
  labs(x = "Chromosome") +
  theme_classic()+
  labs(y=expression(-Log10(q-value))) +
  scale_shape_manual(values=c(21,19,20))+
  scale_color_manual(values=c("grey", "dark grey"), guide =FALSE) 

b3.h <- subset(total.sig,  BLOCK == "3" & Sig.Hi == TRUE)
b3.c <- subset(B3.pv, QCon < 0.1)

B3.pp <- B3.pv %>%
  mutate(
    Group = case_when(
      SNP %in% b3.c$SNP ~ "Control Treatment Filtered",
      SNP %in% all_con$SNP ~ "Control Treatment Filtered",
      SNP %in% con_filter$SNP ~ "Control Treatment Filtered",
      TRUE ~ "Neutral"
    )
  )

man <-ggman(B3.pp, pvalue="QHi", chr="CHR", snp="SNP",bp="BP", relative.positions = TRUE ,sigLine = NA, pointSize=1.5,title = "", alpha = 0.5)

dfm <- man[[1]]

dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))


dfmo <- semi_join(dfm, b3.h, by = "SNP")
dfmo$Group <- "Block 3 Outlier"

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

b3.Hi <-ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(aes(alpha=0.5, shape=Group, size = marker))+  
  #geom_point(aes(x=index, y=log10(QCON), colour = as.factor(chrom_alt), alpha=0.5, shape = Group))+  
  geom_point(data=dfmo, fill="#009E73", aes(x=index, y=marker, shape=Group, size=marker),  color = "#009E73")+
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = FALSE,alpha=FALSE, fill = FALSE, size=FALSE) +
  scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm$marker+1,na.rm=TRUE)))+
  scale_size_continuous(range = c(0.25,3)) +
  labs(x = "Chromosome") +
  ggtitle("Block B") +
  theme_classic()+
  labs(y=expression(-Log10(q-value))) +
  scale_shape_manual(values=c(21,19,20))+
  scale_color_manual(values=c("grey", "dark grey"), guide =FALSE) 


b4.h <- subset(total.sig,  BLOCK == "4" & Sig.Hi == TRUE)
b4.c <- subset(B4.pv, QCon < 0.1)

B4.pp <- B4.pv %>%
  mutate(
    Group = case_when(
      SNP %in% b4.c$SNP ~ "Control Treatment Filtered",
      SNP %in% all_con$SNP ~ "Control Treatment Filtered",
      SNP %in% con_filter$SNP ~ "Control Treatment Filtered",
      TRUE ~ "Neutral"
    )
  )

man <-ggman(B4.pp, pvalue="QHi", chr="CHR", snp="SNP",bp="BP", relative.positions = TRUE ,sigLine = NA, pointSize=1.5,title = "", alpha = 0.5)

dfm <- man[[1]]

dfm$chrom_alt <- factor(dfm$chrom_alt, levels=c(0,1))

dfmo <- semi_join(dfm, b4.h, by = "SNP")
dfmo$Group <- "Block 4 Outlier"

dfmsplit <- split(dfm, dfm$chrom)
xbreaks <- sapply(dfmsplit,function(x) {
    midpoint <- length(x$index)/2
    if(midpoint <1) midpoint <- 1
    return(x$index[midpoint])
})

b4.Hi <-ggplot(dfm, aes(x= index, y=marker, colour = as.factor(chrom_alt)))+
  geom_point(aes(alpha=0.5, shape=Group, size = marker))+  
  geom_point(data=dfmo, fill="#0072B2", aes(x=index, y=marker, shape=Group, size=marker),  color = "#0072B2")+
  scale_x_continuous(breaks = xbreaks, labels = names(xbreaks),expand = c(0,0),limits = c(0,max(dfm$index)+10)) +
  guides(colour = FALSE,alpha=FALSE, fill = FALSE, size=FALSE) +
  #scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm$marker+1,na.rm=TRUE)))+
  scale_y_continuous(expand = c(0,0),limits=c(0,max(dfm$marker+1,na.rm=TRUE)))+
  labs(x = "Chromosome") +
  ggtitle("Block C") +
  scale_size_continuous(range = c(0.25,3)) +
  theme_classic()+
  labs(y=expression(-Log10(q-value))) +
  scale_shape_manual(values=c(21,19,20))+
  scale_color_manual(values=c("grey", "dark grey"), guide =FALSE) 

png(filename="RISG_manhattanplot.png", type="cairo",units="px", width=5600, height=3000, res=300, bg="transparent")

(b2.Hi / b3.Hi / b4.Hi )  + plot_annotation(tag_levels = 'A') + plot_layout(guides = 'collect') 
#(b10.CASE / b11.CASE /b12.CASE )  + plot_annotation(tag_levels = 'A') & theme_black()
#man+scale_fill_manual(values = c("#009E73","#E69F00", "#0072B2"),name = "Legend")
null <- dev.off()
```

## Shared loci across Blocks 

```{bash, eval=FALSE}
cut -f1,2 B2.total.outlier.input | sort -k1,2 | uniq > B2.outlier.loci

cut -f1,2 B3.total.outlier.input | sort -k1,2 | uniq > B3.outlier.loci

cut -f1,2 B4.total.outlier.input | sort -k1,2 | uniq > B4.outlier.loci
```

```{bash, eval=FALSE}
comm -12 <(sort B2.outlier.loci) <(sort B3.outlier.loci) | comm -12 - <(sort B4.outlier.loci) > AB.outlier.loci
```

1108 overlapping loci 

```{bash, eval=FALSE}
vcftools --gzvcf RISGTRSdp20g75nDNA.FIL.recode.vcf.gz --positions AB.outlier.loci --recode --recode-INFO-all --out shared.outlier.RISG.dp20

python2 VCFtoPopPool.py shared.outlier.RISG.dp20.recode.vcf shared.outlier.sync

mawk '!/CHR/' shared.outlier.sync > shared.outlier.input
```

```{bash,eval=FALSE}
./assessPool/scripts/p2/snp-frequency-diff.pl --input shared.outlier.input --min-count 1 --min-coverage 1 --output-prefix shared.outlier --max-coverage 50000 
```

```{bash, eval=FALSE}
mawk -f ./scripts/polarize_freqs shared.outlier_rc | cut -f10-39 | mawk '!/maa/' > shared.outlier.RISG.dp20.pool
```

```{r}
pool.data.shared <- read.table("shared.outlier.RISG.dp20.pool")

df.pool.shared <- apply(pool.data.shared, c(1, 2), function(x) eval(parse(text = x)))

pool.data2.shared <- t(df.pool.shared)

filename.shared <- read.pcadapt(pool.data2.shared, type = "pool")
res.shared <- pcadapt(filename.shared, min.maf = 0.1)
```

```{r}
par(mfrow = c(2, 2))
for (i in 1:4)
  plot(res.shared$loadings[, i], pch = 19, cex = .3, ylab = paste0("Loadings PC", i))
```

```{r}
plot(res.shared,option="screeplot")
```

```{r}
res.shared <- pcadapt(filename.shared, K =4,min.maf = 0.001)
poplist.names <- c("Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress","Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress","Initial","Initial","Initial","Initial","Control","Control","Control","Stress","Stress","Stress")
p1.shared <- plot(res.shared, option = "scores", i = 1, j = 2, pop = poplist.names,
           col = c("#56B4E9", "purple", "orange"))
```

```{r}
poplist.names <- c(rep("Block2", 10),rep("Block3", 10),rep("Block4", 10))
p1.shared <- plot(res.shared, option = "scores", i = 1, j = 2, pop = poplist.names)
```

## Shared genes across Blocks 

```{bash, eval=FALSE}
# For each Block, saving the significant loci as a bed file and then finding genes those loci fall in
cut -f1,2 B2.total.outlier.input | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RISG.Block2.Significant.loci.bed

bedtools intersect -wb -a RISG.Block2.Significant.loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.gene.B2.LOC

cut -f1,2 B3.total.outlier.input | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RISG.Block3.Significant.loci.bed

bedtools intersect -wb -a RISG.Block3.Significant.loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.gene.B3.LOC

cut -f1,2 B4.total.outlier.input | sort -k1,2 | uniq | mawk '{print $1 "\t" $2-1 "\t" $2}' | bedtools sort -i - > RISG.Block4.Significant.loci.bed

bedtools intersect -wb -a RISG.Block4.Significant.loci.bed -b sorted.ref3.0.gene.bed | grep -oh "gene=LOC.*;g" | sed 's/gene=//g' | sed 's/;g//g' | sort | uniq > Sig.gene.B4.LOC
```

```{bash, eval=FALSE}
# Which genes overlap for all Blocks - can reveal shared signatures of selection across distinct genomic pools
comm -12 <(sort Sig.gene.B2.LOC) <(sort Sig.gene.B3.LOC) | comm -12 - <(sort Sig.gene.B4.LOC) > Sig.gene.shared.AB.LOC
```

## Gene Ontology with TopGO

I have a gaf from NCBI with the GO terms for genes in the eastern oyster genome. I'm going to take this file and extract the Gene IDs and associated GO terms.

In terminal:

```{bash}
wget https://ftp.ncbi.nlm.nih.gov/genomes/all/GCF/002/022/765/GCF_002022765.2_C_virginica-3.0/GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

gunzip GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf.gz

mv GCF_002022765.2_C_virginica-3.0_gene_ontology.gaf gene_ontology.gaf
```

Making a new .tab file with just the gene ID and GO IDs (3rd and 5th columns) renamed "gene_name" and "GeneOntologyIDs". The gene_name column will have one unique observation per row and all associated GO terms will be saved in the same row, with values separaetd by ;

```{bash}
awk 'NR>9 {print $3 "\t" $5}' gene_ontology.gaf | 
sed '1i gene_name\tGeneOntologyIDs' | 
sort | awk '
NR==1 {print; next}
$1!=prev {
if (NR>2) print prev "\t" goids;
prev=$1;
goids=$2;
next
}
{
goids=goids "; " $2
}
END {print prev "\t" goids}
' > gene2go_full.tab
```

Gene IDs for the candidate loci are saved to `Sig.loci.TOTAL.RISG.LOC`

I'm making another file with the gene IDs and GO IDs for the candidate loci:

```{bash}
grep -F -f Sig.loci.TOTAL.RISG.LOC gene2go_full.tab > risg_ab_cand_genes_GO.tab

# Add header
awk 'BEGIN {print "gene_name\tGeneOntologyIDs"} {print}' risg_ab_cand_genes_GO.tab > risg_ab_cand_genes_GO_head.tab
```

In R:

```{r}
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("topGO")
```

```{r setup, include=FALSE}
library(topGO)
```

Read in gene2go information 

Old file names: all_genes (gene2go_full.tab); candidate genes (gene2go.tab)

```{r}
all_gene2go<-read.delim("gene2go_full.tab", sep="\t")
out_gene2go<-read.delim("risg_ab_cand_genes_GO_head.tab", sep="\t")
```

Make list of genes for input to TopGO

```{r}
cand_genes <- as.character(unique(out_gene2go$gene_name))
all_genes <- as.character(all_gene2go$gene_name)
out_GeneList <- factor(as.integer(all_genes %in% cand_genes))
names(out_GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test  

Read in gene-to-go-mappings
```{r}
gene2go_topgo<-readMappings("gene2go_full.tab", IDsep=";", sep="\t")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_BP_pi <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_BP_FE_pi <- runTest(GO_OUT_BP_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_BP_En_pi <- GenTable(GO_OUT_BP_pi, Fisher = GO_OUT_BP_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_BP_En_pi$Fisher<-as.numeric(GO_OUT_BP_En_pi$Fisher)
GO_OUT_BP_En_sig_pi<-GO_OUT_BP_En_pi[GO_OUT_BP_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_BP_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_BP_En_sig_pi$GO.ID <- trimws(GO_OUT_BP_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_BP_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_BP_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_BP_En_sig_gene_pi$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_CC_pi <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_CC_FE_pi <- runTest(GO_OUT_CC_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_CC_En_pi <- GenTable(GO_OUT_CC_pi, Fisher = GO_OUT_CC_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_CC_En_pi$Fisher<-as.numeric(GO_OUT_CC_En_pi$Fisher)
GO_OUT_CC_En_sig_pi<-GO_OUT_CC_En_pi[GO_OUT_CC_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_CC_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_CC_En_sig_pi$GO.ID <- trimws(GO_OUT_CC_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_CC_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_CC_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_CC_En_sig_gene_pi$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_MF_pi <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_MF_FE_pi <- runTest(GO_OUT_MF_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_MF_En_pi <- GenTable(GO_OUT_MF_pi, Fisher = GO_OUT_MF_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_MF_En_pi$Fisher<-as.numeric(GO_OUT_MF_En_pi$Fisher)
GO_OUT_MF_En_sig_pi<-GO_OUT_MF_En_pi[GO_OUT_MF_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_MF_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_MF_En_sig_pi$GO.ID <- trimws(GO_OUT_MF_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_MF_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_MF_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_MF_En_sig_gene_pi$ontology <- "Molecular Functions"
```

### Join ontologies

Bind `GO_OUT_BP_En_sig_gene_pi`, `GO_OUT_CC_En_sig_gene_pi`, and `GO_OUT_MF_En_sig_gene_pi` to make a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_OUT_En_sig_gene_pi <- rbind(GO_OUT_BP_En_sig_gene_pi, GO_OUT_CC_En_sig_gene_pi, GO_OUT_MF_En_sig_gene_pi)
```

Merge `pi_bed_filt` with `GO_PEVE_En_sig_gene` and calculate proportion of significant genes
```{r}
merged_out_genes <- GO_OUT_En_sig_gene_pi %>%
  mutate(prop.sig.genes = Significant/Annotated)

# Write to csv 
write.csv(merged_out_genes, "RISG_OUT_AB_GO_en_sig_gene.csv")
```

Plot  
```{r}
GO_plot_pi<-ggplot(merged_out_genes, aes(x = Term, y = -log(Fisher), size = prop.sig.genes, fill = -log(Fisher))) +
  expand_limits(y = 1.5) +
  ylim(1, 7.25) +
  #Add horizontal lines with a single aesthetic value
  geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); GO_plot_pi
```



### Repeating just for genes overlapped by all Blocks

Gene IDs for the shared candidate loci for all blocks are saved to `Sig.gene.shared.AB.LOC`

I'm making another file with the gene IDs and GO IDs for the candidate loci:

```{bash}
grep -F -f Sig.gene.shared.AB.LOC gene2go_full.tab > risg_shared_cand_genes_GO.tab

# Add header
awk 'BEGIN {print "gene_name\tGeneOntologyIDs"} {print}' risg_shared_cand_genes_GO.tab > risg_shared_cand_genes_GO_head.tab
```

In R:

```{r}
#if (!require("BiocManager", quietly = TRUE))
    #install.packages("BiocManager")

#BiocManager::install("topGO")
```

```{r setup, include=FALSE}
library(topGO)
```

Read in gene2go information 

Old file names: all_genes (gene2go_full.tab); candidate genes (gene2go.tab)

```{r}
all_gene2go<-read.delim("gene2go_full.tab", sep="\t")
out_gene2go<-read.delim("risg_shared_cand_genes_GO_head.tab", sep="\t")
```

Make list of genes for input to TopGO

```{r}
cand_genes <- as.character(unique(out_gene2go$gene_name))
all_genes <- as.character(all_gene2go$gene_name)
out_GeneList <- factor(as.integer(all_genes %in% cand_genes))
names(out_GeneList) <- all_genes
```

The following code will perform GO enrichment using the weighted Fisher's exact test  

Read in gene-to-go-mappings
```{r}
gene2go_topgo<-readMappings("gene2go_full.tab", IDsep=";", sep="\t")
```

Set function to select genes of interest (ie those that have pvalue < 0.05)
```{r}
topDiffGenes <- function(allScore) {
return(allScore < 0.05)}
```

### Biological Processes

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_BP_pi <-new("topGOdata", ontology="BP", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_BP_FE_pi <- runTest(GO_OUT_BP_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_BP_En_pi <- GenTable(GO_OUT_BP_pi, Fisher = GO_OUT_BP_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_BP_En_pi$Fisher<-as.numeric(GO_OUT_BP_En_pi$Fisher)
GO_OUT_BP_En_sig_pi<-GO_OUT_BP_En_pi[GO_OUT_BP_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_BP_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_BP_En_sig_pi$GO.ID <- trimws(GO_OUT_BP_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_BP_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_BP_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_BP_En_sig_gene_pi$ontology <- "Biological Processes"
```

### Cellular Components

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_CC_pi <-new("topGOdata", ontology="CC", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_CC_FE_pi <- runTest(GO_OUT_CC_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_CC_En_pi <- GenTable(GO_OUT_CC_pi, Fisher = GO_OUT_CC_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_CC_En_pi$Fisher<-as.numeric(GO_OUT_CC_En_pi$Fisher)
GO_OUT_CC_En_sig_pi<-GO_OUT_CC_En_pi[GO_OUT_CC_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_CC_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_CC_En_sig_pi$GO.ID <- trimws(GO_OUT_CC_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_CC_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_CC_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_CC_En_sig_gene_pi$ontology <- "Cellular Components"
```

### Molecular Functions

Create `topGOdata` object, which is required for topGO analysis
```{r}
GO_OUT_MF_pi <-new("topGOdata", ontology="MF", gene2GO=gene2go_topgo, allGenes=out_GeneList, annot = annFUN.gene2GO, geneSel=topDiffGenes)
```

Run GO enrichment test 
```{r}
GO_OUT_MF_FE_pi <- runTest(GO_OUT_MF_pi, algorithm="weight01", statistic="fisher")
```

Generate results table 
```{r}
GO_OUT_MF_En_pi <- GenTable(GO_OUT_MF_pi, Fisher = GO_OUT_MF_FE_pi, orderBy = "Fisher",  topNodes = 100, numChar = 100)
```
Only taking the top 100 GO terms 

Filter by significant results
```{r}
GO_OUT_MF_En_pi$Fisher<-as.numeric(GO_OUT_MF_En_pi$Fisher)
GO_OUT_MF_En_sig_pi<-GO_OUT_MF_En_pi[GO_OUT_MF_En_pi$Fisher<0.05,]
```

Merge `GO_OUT_MF_En_sig_pi` with GO and gene info. 
```{r}
# Separate GO terms 
out_gene2go <- out_gene2go %>%
  separate_rows(GeneOntologyIDs, sep = ";")

# Ensure GO terms in both datasets are formatted similarly (trim whitespaces)
out_gene2go$GeneOntologyIDs <- trimws(out_gene2go$GeneOntologyIDs)
GO_OUT_MF_En_sig_pi$GO.ID <- trimws(GO_OUT_MF_En_sig_pi$GO.ID)

# Join the datasets based on GO term
GO_OUT_MF_En_sig_gene_pi <- out_gene2go %>%
  left_join(GO_OUT_MF_En_sig_pi, by = c("GeneOntologyIDs" = "GO.ID")) %>%
  na.omit()

# Add ontology column 
GO_OUT_MF_En_sig_gene_pi$ontology <- "Molecular Functions"
```

### Join ontologies

Bind `GO_OUT_BP_En_sig_gene_pi`, `GO_OUT_CC_En_sig_gene_pi`, and `GO_OUT_MF_En_sig_gene_pi` to make a df that has significantly enriched GO terms for all ontologies 
```{r}
GO_OUT_En_sig_gene_pi <- rbind(GO_OUT_BP_En_sig_gene_pi, GO_OUT_CC_En_sig_gene_pi, GO_OUT_MF_En_sig_gene_pi)
```

Merge `pi_bed_filt` with `GO_PEVE_En_sig_gene` and calculate proportion of significant genes
```{r}
merged_out_genes <- GO_OUT_En_sig_gene_pi %>%
  mutate(prop.sig.genes = Significant/Annotated)

# Write to csv 
write.csv(merged_out_genes, "RISG_OUT_SHARED_GO_en_sig_gene.csv")
```

Plot  
```{r}
GO_plot_pi<-ggplot(merged_out_genes, aes(x = Term, y = -log(Fisher), size = prop.sig.genes, fill = -log(Fisher))) +
  expand_limits(y = 1.5) +
  ylim(1, 7.25) +
  #Add horizontal lines with a single aesthetic value
  geom_hline(yintercept = -log10(0.01), linetype = "longdash", colour = "black", linewidth = .6) +
  geom_hline(yintercept = -log10(0.001), linetype = "solid", colour = "black", linewidth = .6) +
  geom_point(shape = 21) + 
  scale_size(range = c(2, 12)) + 
  scale_fill_continuous(low = "#1AD3D1FF", high = "#4686FBFF") +
  xlab('') + 
  ylab('Enrichment score') +
  #labs(caption = 'Cut-off lines at p=0.01 and p=0.001') +
  theme_bw() +
  facet_grid(vars(ontology), scales = "free", space = "free_y") +
  coord_flip(); GO_plot_pi
```

## Initial timepoint nucleotide diversity

Using grenedalf:

[grenedalf Github](https://github.com/lczech/grenedalf/wiki/Subcommand:-diversity#description)

In PATH: `/home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering`

```{bash}
mamba env create -c conda-forge -c bioconda -n grenedalf grenedalf
conda activate grenedalf
```

```{bash}
mkdir grenedalf
cd grenedalf
```

```{bash}
ln -s ../SNP.RISG.TRSdp.20.B90.2a.perp.vcf.gz .
ln -s ../reference.fasta .
ln -s ../reference.fasta.fai .
ln -s ../T0_samples .
```

I needed switch the order of the chromosomes in the .fasta.fai file so that NC_035789.1 was at the end. 

```
NC_007175.2   17244   13    17244   17245
NC_035780.1   65668440    32667330    65668440    65668441
NC_035781.1   61752955    98335784    61752955    61752956
NC_035782.1   77061148    160088753   77061148    77061149
NC_035783.1   59691872    237149915   59691872    59691873
NC_035784.1   98698416    296841801   98698416    98698417
NC_035785.1   51258098    395540231   51258098    51258099
NC_035786.1   57830854    446798343   57830854    57830855
NC_035787.1   75944018    504629211   75944018    75944019
NC_035788.1   104168038   580573243   104168038   104168039
NC_035789.1   32650045    17271   32650045    32650046
```

```{bash}
grenedalf diversity \
    --vcf-path /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/SNP.RISG.TRSdp.20.B90.2a.perp.vcf.gz \
    --reference-genome-fai reference.fasta.fai \
    --filter-samples-include T0_samples \
    --pool-sizes 80000 \
    --window-type interval \
    --window-interval-width 10000 \
    --window-interval-stride 0 \
    --window-average-policy valid-loci \
    --filter-sample-min-count 2 \
    --filter-sample-min-read-depth 10
```

For most of the inputs, I used the default options. For --pool-sizes, it's the number of individuals in the sample x2 as oysters are diploid. 

```
# Output

At chromosome NC_035780.1
At chromosome NC_035781.1
At chromosome NC_035782.1
At chromosome NC_035783.1
At chromosome NC_035784.1
At chromosome NC_035785.1
At chromosome NC_035786.1
At chromosome NC_035787.1
At chromosome NC_035788.1
At chromosome NC_035789.1

Processed 68108 windows.
Processed 10 chromosomes with 980787 positions.

Sample filter summary (summed up across all samples):
Missing: 194959
Passed:  11574485

Total filter summary (after applying all sample filters):
Invariant: 5383
Passed:    975404

Finished 2025-04-18 17:38:07
```

This produces a .csv file called diversity .csv

I made a new directory `T0_diversity` and moved the .csv to this location for organizational purposes. I also renamed the csv to T0_diversity.csv.

Read in .csv file

```{r}
larv.nuc.diversity <- read.csv("/home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/T0_diversity/T0_diversity.csv")
```

I need to modify this dataframe quite a bit to be able to run statistical analyses:

```{r}
library(dplyr)
library(tidyr)
library(stringr)
```

**Pivot to Long Format**
You want to gather all the population-specific columns (like B2_T0_1.theta_pi, etc.) into long format.

```{r}
# Identify columns for pivoting
pop_cols <- grep("B[234]_T0_\\d+\\.(missing|numeric|passed|theta_pi|theta_watterson|tajimas_d)", 
                 names(larv.nuc.diversity), value = TRUE)

larv_long <- larv.nuc.diversity %>%
  pivot_longer(
    cols = all_of(pop_cols),
    names_to = c("POP", "METRIC"),
    names_pattern = "(B[234]_T0_\\d+)\\.(.*)",
    values_to = "VALUE"
  )
```

**Filter for theta_pi and passed**
You want one row per population per bin, with theta_pi as PI and passed as N_VARIANTS.

```{r}
larv_wide <- larv_long %>%
  filter(METRIC %in% c("theta_pi", "passed")) %>%
  pivot_wider(
    names_from = METRIC,
    values_from = VALUE
  )
```

**Add GROUP Column**
Extract group (B2, B3, B4) from POP.

```{r}
larv_final <- larv_wide %>%
  mutate(
    GROUP = str_extract(POP, "B[234]"),
    CHROM = chrom,
    BIN_START = start,
    BIN_END = end,
    N_VARIANTS = as.integer(passed),
    PI = as.numeric(theta_pi)
  ) %>%
  dplyr::select(CHROM, BIN_START, BIN_END, N_VARIANTS, PI, POP, GROUP)
```

**Remove Rows with Missing PI or N_VARIANTS**

```{r}
larv_final <- larv_final %>%
  filter(!is.na(PI), !is.na(N_VARIANTS))
```

```{r}
larv_final$GROUP <- factor(larv_final$GROUP, levels=c("B2", "B3","B4"))
larv_final$POP <- factor(larv_final$POP, levels=c("B2_T0_1", "B2_T0_2", "B2_T0_3", "B2_T0_4", "B3_T0_1", "B3_T0_2", "B3_T0_3", "B3_T0_4", "B4_T0_1", "B4_T0_2", "B4_T0_3", "B4_T0_4"))
```

**Plotting**

```{r}
bd.pi <-ggplot(larv_final, aes(x=PI,y = POP))+
  #geom_violin(aes(color=GROUP,fill= GROUP)) +
  #stat_summary(fun=mean, geom="point", shape=23, size=4)+
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.005)+
  geom_boxplot(aes(fill=GROUP), width=0.75,outlier.shape = 23, outlier.color = "black")+
  geom_vline(xintercept =mean(larv_final$PI),linetype = "dashed") +
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#CC79A7", "#009E73", "#0072B2"))+
  scale_color_manual(values=c("#CC79A7", "#009E73", "#0072B2"))+
  scale_y_discrete(limits = rev(levels(larv_final$POP)))+
  #ylab("Population/Line")+
  labs(x=expression(pi)) +
  ggtitle("Experimental Block Initial TimePoint Nucleotide Diversity")+
  theme_classic() +
  theme(axis.title.y = element_blank()) 
bd.pi
```

**Stats**

```{r}
library(car)
library(MASS)
```

I'm going to run an ANOVA to compare PI between each experimental block.

```{r}
anova_model <- aov(PI ~ GROUP, data = larv_final)
```

I need to check for normality: 

```{r}
qqPlot(residuals(anova_model)) 
```

The data is not normally distributed. I'm going to do a non-parametric test

```{r}
kruskal.test(PI ~ GROUP, data = larv_final)
```

The ANOVA is significant, so I will run a post hoc test.

```{r}
pairwise.wilcox.test(larv_final$PI, larv_final$GROUP, p.adjust.method = "bonferroni")
```

It looks like all groups are significantly different from one another.


I want to double check the significance between groups and check effect sizes. The p-values are really small which may be because there are so many data points. Effect size may allow for a more accurate comparison between groups. First I will split up the data set into sets of two groups:

```{r}
# Create a new data frame excluding one group
B2_B3_PI <-larv_final[larv_final$GROUP != "B4", ]
B2_B4_PI <-larv_final[larv_final$GROUP != "B3", ]
B3_B4_PI <-larv_final[larv_final$GROUP != "B2", ]
```

This allows me to run t-tests using infer

```{r}
library(infer)
library(plotrix)
```

```{r}
B2_B3_PI %>% t_test(formula = PI ~ GROUP, alternative = "two.sided",order = c("B2", "B3"))
```

I'm also going to calculate eta squared: 

```{r}
lm = aov(B2_B3_PI$PI ~ B2_B3_PI$GROUP)
summary(lm)
effect_size <- effectsize::eta_squared(lm, partial=F)
effect_size
```

Repeat for comparison between B2 and B4:

```{r}
B2_B4_PI %>% t_test(formula = PI ~ GROUP, alternative = "two.sided",order = c("B2", "B4"))
```

```{r}
lm = aov(B2_B4_PI$PI ~ B2_B4_PI$GROUP)
summary(lm)
effect_size <- effectsize::eta_squared(lm, partial=F)
effect_size
```

Repeat for comparison between B3 and B4:

```{r}
B3_B4_PI %>% t_test(formula = PI ~ GROUP, alternative = "two.sided",order = c("B3", "B4"))
```

```{r}
lm = aov(B3_B4_PI$PI ~ B3_B4_PI$GROUP)
summary(lm)
effect_size <- effectsize::eta_squared(lm, partial=F)
effect_size
```

So the t-tests confirm the ANOVA results and effect sizes are really low across the board. I think this is the nature of the data. 

**Calculate summary statistics**

```{r}
risg_pi_summary <- larv_final %>%
  group_by(GROUP) %>%
  summarize(
    count = n(),
    min = min(PI, na.rm = TRUE),
    q1 = quantile(PI, 0.25, na.rm = TRUE),
    median = median(PI, na.rm = TRUE),
    mean = mean(PI, na.rm = TRUE),
    q3 = quantile(PI, 0.75, na.rm = TRUE),
    max = max(PI, na.rm = TRUE),
    sd = sd(PI, na.rm = TRUE)
  )
```




### Repeating to compare T0 to TJE within each Block

I made two new files: the `sample_names` file includes all the sample names (may not be necessary since these are all the samples in the vcf file); and the `pool_size` file is all the individuals in the first column with their pool sizes in the second column, tab delimited. Again, pool sizes are the number of individuals in each pool x2 as oysters are diploid. 

```{bash}
head /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/pool_sizes
```

```{bash}
grenedalf diversity \
    --vcf-path /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/SNP.RISG.TRSdp.20.B90.2a.perp.vcf.gz \
    --reference-genome-fai reference.fasta.fai \
    --filter-samples-include sample_names \
    --pool-sizes pool_sizes \
    --window-type interval \
    --window-interval-width 10000 \
    --window-interval-stride 0 \
    --window-average-policy valid-loci \
    --filter-sample-min-count 2 \
    --filter-sample-min-read-depth 10 \
    --out-dir /home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/wBlock_diversity \
    --file-prefix all_samples_
```

```
# Output

Started 2025-07-07 11:09:20

Reading reference genome fai
Reference genome contains 11 chromosomes
Sample filter results in 30 samples out of 30 being used.
Processing 30 samples

At chromosome NC_035780.1
At chromosome NC_035781.1
At chromosome NC_035782.1
At chromosome NC_035783.1
At chromosome NC_035784.1
At chromosome NC_035785.1
At chromosome NC_035786.1
At chromosome NC_035787.1
At chromosome NC_035788.1
At chromosome NC_035789.1

Processed 68108 windows.
Processed 10 chromosomes with 980787 positions.

Sample filter summary (summed up across all samples):
Missing: 2870366
Passed:  26553244

Total filter summary (after applying all sample filters):
Invariant: 89
Passed:    980698

Finished 2025-07-07 11:15:13
```

Read in .csv file

```{r}
larv.nuc.diversity.all <- read.csv("/home/azyck/RISG_Larval/RISG_ddocent/RISG_Larval_SNPFiltering/grenedalf/wBlock_diversity/all_samples_diversity.csv")
```

**Pivot to Long Format**
You want to gather all the population-specific columns (like B2_T0_1.theta_pi, etc.) into long format.

```{r}
pop_cols_all <- grep(
  "B[234]_T(?:0|CE)_(?:\\d+|CB\\d+|HB\\d+)\\.(missing|numeric|passed|theta_pi|theta_watterson|tajimas_d)",
  names(larv.nuc.diversity.all), value = TRUE
)

larv_long_all <- larv.nuc.diversity.all %>%
  pivot_longer(
    cols = all_of(pop_cols_all),
    names_to = c("POP", "METRIC"),
    names_pattern = "(B[234]_T(?:0|CE)_(?:\\d+|CB\\d+|HB\\d+))\\.(.*)",
    values_to = "VALUE"
  )
```

**Filter for theta_pi and passed**
You want one row per population per bin, with theta_pi as PI and passed as N_VARIANTS.

```{r}
larv_wide_all <- larv_long_all %>%
  filter(METRIC %in% c("theta_pi", "passed")) %>%
  pivot_wider(
    names_from = METRIC,
    values_from = VALUE
  )
```

**Add GROUP Column**
Extract group (B2, B3, B4) from POP.

```{r}
larv_final_all <- larv_wide_all %>%
  mutate(
    GROUP = str_extract(POP, "B[234]"),
    TIMEPOINT = str_extract(POP, "T(?:0|CE)"),
    CHROM = chrom,
    BIN_START = start,
    BIN_END = end,
    N_VARIANTS = as.integer(passed),
    PI = as.numeric(theta_pi),
    TREATMENT = case_when(
      str_detect(POP, "T0") ~ "T0",
      str_detect(POP, "CB\\d+$") ~ "CB",
      str_detect(POP, "HB\\d+$") ~ "HB",
      TRUE ~ NA_character_
    )
  ) %>%
  dplyr::select(CHROM, BIN_START, BIN_END, N_VARIANTS, PI, POP, GROUP, TIMEPOINT, TREATMENT)
```

**Remove Rows with Missing PI or N_VARIANTS**

```{r}
larv_final_all <- larv_final_all %>%
  filter(!is.na(PI), !is.na(N_VARIANTS))
```

```{r}
larv_final_all$GROUP <- factor(larv_final_all$GROUP, levels=c("B2", "B3","B4"))
larv_final_all$TIMEPOINT <- factor(larv_final_all$TIMEPOINT, levels=c("T0", "TCE"))
larv_final_all$TREATMENT <- factor(larv_final_all$TREATMENT, levels=c("T0", "CB","HB"))
larv_final_all$POP <- factor(larv_final_all$POP, levels=c("B2_T0_1", "B2_T0_2", "B2_T0_3", "B2_T0_4", "B2_TCE_CB1", "B2_TCE_CB2", "B2_TCE_CB3", "B2_TCE_HB1", "B2_TCE_HB2", "B2_TCE_HB3", "B3_T0_1", "B3_T0_2", "B3_T0_3", "B3_T0_4", "B3_TCE_CB1", "B3_TCE_CB2", "B3_TCE_CB3", "B3_TCE_HB1", "B3_TCE_HB2", "B3_TCE_HB3", "B4_T0_1", "B4_T0_2", "B4_T0_3", "B4_T0_4", "B4_TCE_CB1", "B4_TCE_CB2", "B4_TCE_CB3", "B4_TCE_HB1", "B4_TCE_HB2", "B4_TCE_HB3"))
```

**Plotting**

```{r}
bd.pi.all <-ggplot(larv_final_all, aes(x=PI,y = POP))+
  #geom_violin(aes(color=GROUP,fill= GROUP)) +
  #stat_summary(fun=mean, geom="point", shape=23, size=4)+
  #geom_dotplot(binaxis='y', stackdir='center', dotsize=0.005)+
  geom_boxplot(aes(fill=GROUP), width=0.75,outlier.shape = 23, outlier.color = "black")+
  geom_vline(xintercept =mean(larv_final_all$PI),linetype = "dashed") +
  stat_summary(fun=mean, geom="point", shape=23, size=2)+
  scale_fill_manual(values=c("#CC79A7", "#009E73", "#0072B2"))+
  scale_color_manual(values=c("#CC79A7", "#009E73", "#0072B2"))+
  scale_y_discrete(limits = rev(levels(larv_final_all$POP)))+
  #ylab("Population/Line")+
  labs(x=expression(pi)) +
  ggtitle("Nucleotide Diversity")+
  theme_classic() +
  theme(axis.title.y = element_blank()) 
bd.pi.all
```

**Stats**

First, I need to subset the meta data frame by Block:

```{r}
block2_df <- filter(larv_final_all, GROUP == "B2")
block3_df <- filter(larv_final_all, GROUP == "B3")
block4_df <- filter(larv_final_all, GROUP == "B4")
```

I'm going to run an ANOVA to compare PI between T0 and the the final timepoint samples CB and HB within each experimental block.

**Block 2**

```{r}
b2_aov_model <- aov(PI ~ TREATMENT, data = block2_df)
```

I need to check for normality: 

```{r}
qqPlot(residuals(b2_aov_model)) 
```

The data is not normally distributed. I'm going to do a non-parametric test

```{r}
kruskal.test(PI ~ TREATMENT, data = block2_df)
```

Also running separate t-tests comparing PI between T0 and CB 

```{r}
b2_t0_cb <- subset(block2_df, TREATMENT %in% c("T0", "CB"))
t.test(PI ~ TREATMENT, data = b2_t0_cb)
```

and T0 and HB

```{r}
b2_t0_hb <- subset(block2_df, TREATMENT %in% c("T0", "HB"))
t.test(PI ~ TREATMENT, data = b2_t0_hb)
```


**Block 3**

```{r}
b3_aov_model <- aov(PI ~ TREATMENT, data = block3_df)
```

I need to check for normality: 

```{r}
qqPlot(residuals(b3_aov_model)) 
```

The data is not normally distributed. I'm going to do a non-parametric test

```{r}
kruskal.test(PI ~ TREATMENT, data = block3_df)
```

The ANOVA is significant, so I will run a post hoc test.

```{r}
pairwise.wilcox.test(block3_df$PI, block3_df$TREATMENT, p.adjust.method = "bonferroni")
```

Also running separate t-tests comparing PI between T0 and CB 

```{r}
b3_t0_cb <- subset(block3_df, TREATMENT %in% c("T0", "CB"))
t.test(PI ~ TREATMENT, data = b3_t0_cb)
```

and T0 and HB

```{r}
b3_t0_hb <- subset(block3_df, TREATMENT %in% c("T0", "HB"))
t.test(PI ~ TREATMENT, data = b3_t0_hb)
```

**Block 4**

```{r}
b4_aov_model <- aov(PI ~ TREATMENT, data = block4_df)
```

I need to check for normality: 

```{r}
qqPlot(residuals(b4_aov_model)) 
```

The data is not normally distributed. I'm going to do a non-parametric test

```{r}
kruskal.test(PI ~ TREATMENT, data = block4_df)
```

The ANOVA is significant, so I will run a post hoc test.

```{r}
pairwise.wilcox.test(block4_df$PI, block4_df$TREATMENT, p.adjust.method = "bonferroni")
```

Also running separate t-tests comparing PI between T0 and CB 

```{r}
b4_t0_cb <- subset(block4_df, TREATMENT %in% c("T0", "CB"))
t.test(PI ~ TREATMENT, data = b4_t0_cb)
```

and T0 and HB

```{r}
b4_t0_hb <- subset(block4_df, TREATMENT %in% c("T0", "HB"))
t.test(PI ~ TREATMENT, data = b4_t0_hb)
```

To summarize: for Block 2, there is not a significant difference in genetic diversity from initial to end of exposure. For Block 3, there is a significant difference between initial and end for both Control and Stress treatments. For Block 4, there is a significant difference between initial and end for Stress but not Control. 

**Calculate summary statistics**

```{r}
risg_pi_summary_all <- larv_final_all %>%
  group_by(GROUP, TREATMENT) %>%
  summarize(
    count = n(),
    min = min(PI, na.rm = TRUE),
    q1 = quantile(PI, 0.25, na.rm = TRUE),
    median = median(PI, na.rm = TRUE),
    mean = mean(PI, na.rm = TRUE),
    q3 = quantile(PI, 0.75, na.rm = TRUE),
    max = max(PI, na.rm = TRUE),
    sd = sd(PI, na.rm = TRUE)
  )
```

